### 2025-09-21 Manuscript_MD_ScoreReLiability.qmd
update manuscript up to appendix

---
title: "Manuscript"
bibliography: ReLiability_references.bib
csl: styles/apa-6th-edition.csl
format: pdf
editor: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
# library loading and installing as necessary

# relevant libraries required for this script
packages <- c("ggplot2", "here", "metafor", "gridExtra", "knitr", "magrittr", "dplyr", "gt")

# check, whether library already installed or not - install and load as needed:
apply(as.matrix(packages), MARGIN = 1, FUN = function(x) {

  pkg_avail <- nzchar(system.file(package = x))   # check if library is installed on system

  if(pkg_avail){
    require(x, character.only = TRUE)             # load the library, if already installed

  }else{
    install.packages(x, repos = "https://ftp.gwdg.de/pub/misc/cran/")                           # install the library, if missing
    require(x, character.only = TRUE)             # load after installation
  }
})

source(here("ReLiability_Function-library.R"))

ES_rma_df <- read.csv(here("Data/Processed/Aggregates_ES_analysis.csv"))

agg_L <- readRDS(here("Data/Processed/Aggregates_simple.csv"))

MASC_names <- unique(ES_rma_df$MASC)

```

# Liability or reliability? Exploring the role of effect size attenuation on meta-analytic heterogeneity

Lukas J. Beinhauer *University of Erfurt* ORCID:
https://orcid.org/0000-0001-9841-3089

Jens H. Fünderich *Ludwig Maximilian University of Munich* ORCID:
https://orcid.org/0000-0002-7185-9248

Frank Renkewitz *University of Erfurt* ORCID:
https://orcid.org/0000-0001-8072-6802

Large-scale collaborative replication efforts have sparked discussions
surrounding the replicability of psychological phenomena. For attempts
to estimate a phenomenon’s replicability or to predict whether a future
replication will replicate successfully, heterogeneity is a crucial
parameter that needs to be assessed [@stanley2018]. In the meta-analytic
context, heterogeneity describes the variation of effect sizes, free of
sampling error [@dersimonian1986; @higgins2002; @riley2011]. Therefore,
the presence of heterogeneity can imply that some replications of a
single phenomenon may be successful, while others are not. If
heterogeneity increases, meaning the phenomenon’s effect size varies
more strongly for unexplained reasons, the probability of observing an
effect size around zero or even in the negative space grows larger as
well [@kenny2019]. If we know the mean size and heterogeneity of a
phenomenon’s effect, we would theoretically be able to establish an
expected replication rate [e.g. @vu2024].

Similarly, it has been argued that heterogeneity in effect sizes is an
indicator of the theory’s “completeness” surrounding the phenomenon. For
example, Linden and Hönekopp argue that “low (as opposed to high)
heterogeneity reflects a more advanced understanding of the subject
matter being studied” [-@LindenHönekopp, p. 2]. Similarly, von Hippel
and Schuetze [-@vonHippel2023] argue that heterogeneity in effects is an
indicator of a vague, poorly specified theory.

Large-scale attempts of direct replications, using identical protocols,
such as the Many Labs studies or Registered Replication Reports
[@ManyLabs1; e.g. @RRRHart], for the first time allow researchers to
estimate heterogeneity undistorted by selection processes like
publication bias. In re-analyses of these studies, Olsson-Collentine et
al. [-@olsson2020] identify a positive correlation between a
phenomenon’s effect size and its heterogeneity. Both van Erp et al.
[-@vanErp2017] and Stanley et al. [-@stanley2018] estimate strong
degrees of heterogeneity across conceptual replications in Psychology.
In a separate re-analysis of large scale direct replications, Renkewitz
et al. (in preparation) identify substantial heterogeneity in almost all
projects where a non-zero effect could be identified. This aligns with
the correlation found by Olsson-Collentine et al. [-@olsson2020].

One source for this heterogeneity, which is not a typical moderator tied
to a specific theory, might be impairments in score reliability which
vary in their extent across replications. \[This is an argument put
forward repeatedly by original researchers and replicators alike, such
as Olsson-Collentine et al. [-@olsson2020] (do we know any more?).\]
Hunter and Schmidt [-@hunter2004] and Wiernik and Dahlke
[-@wiernik2020], discussing the role of score reliability for
standardized effect sizes, claim that differences in score reliability
should inflate, and therefore increase, heterogeneity. This implies that
standard corrections for imperfect score reliability, such as
attenuation correction procedures, should remove that inflation,
reducing heterogeneity in corrected effect sizes.

Over the following pages, we assess whether score reliability can in
fact be made responsible for heterogeneity identified across several
psychological phenomena. However, we find no empirical evidence for the
argument that correcting for differences in score reliability reduces
effect size heterogeneity. Instead, to our great surprise, we find that
heterogeneity might as well be larger after the correction procedure.
Therefore, previous analyses and assumptions regarding the relationship
between score reliability and heterogeneity need to be false. In the
second half of this text we make use of analytical arguments to identify
the necessary conditions under which an attenuation correction
inevitably leads to an increase in effect size heterogeneity \[Do we
need any "Schleifen" in this paragraph - to reduce informational load /
density\].

## Effect sizes and score reliability in meta-analysis

In both the initial reports of large scale replication attempts, as well
as the re-analyses by Olsson-Collentine et al. [-@olsson2020],
standardized effect sizes (from here-on abbreviated as ES) such as
Cohen’s d or Hedge’s g were used. For the remainder of the article,
Cohen’s d, defined in @eq-d, will be used as an exemplary estimate of
ES, as it is used across a wide range of contexts and well understood by
a broad audience.

$$d = \frac{MD}{\sigma_X}$$ {#eq-d}

Here, MD refers to the mean difference between two groups of interest,
while $\sigma_X$ is the pooled standard deviation. As is widely known,
score reliability affects ES [e.g. @hunter2004; @wiernik2020]. In the
context of classical test theory (CTT) score reliability $\rho_{XX’}$ is
defined as the ratio of true to total score variance [@lord2008], as
defined in @eq-rel.

$$\rho_{XX’} = \frac{\sigma^2_T}{\sigma^2_X}=\frac{\sigma^2_T}{\sigma^2_T + \sigma^2_E}$$ {#eq-rel}

In this equation $\sigma_T^2$ refers to the true score variance in the
sense of CTT, meaning the actual variance of the variable, undistorted
by random measurement error. In the same sense, $\sigma_X^2$ refers to
the total variance of scores, including both the true variance and the
random error variance $\sigma_E^2$. This implies that, if true score
variance is assumed to be constant across replications, a lower score
reliability can only occur due to a larger error score variance. Total
score variance would also be larger, therein leading to a smaller ES, as
opposed to a similar data-set where a higher score reliability is
achieved through a smaller error score variance. However, score
reliability is an aspect of a measuring instrument applied to a
population [@lord2008; @vacha1998]. In the meta-analytic context, as
studies tend to be replicated in a variety of populations, neither true
nor error score variance can be expected to always be identical across
replications. This most likely leads to heterogeneity in score
reliability, which, as discussed, is bound to affect ES heterogeneity as
well.

Previous discussions of heterogeneity in score reliability have
exclusively discussed it as a parameter that inflates heterogeneity in
ES, implying that, if score reliability was perfect across all
replications, the actual heterogeneity would have been lower. In their
discussion, Wiernik and Dahlke claim that „*Measurement error variance
will impact the results of meta-analyses in three ways: by (a) biasing
the mean effect size toward zero, (b) inflating effect-size
heterogeneity and confounding moderator effects, and (c) confounding
publication-bias and sensitivity analyses*” [p. 3, -@wiernik2020].
Additionally, more clearly, they state that “*If the studies included in
a meta-analysis differ in their measures’ reliabilities, heterogeneity
estimates will be artifactually inflated, erroneously suggesting larger
potential moderator effects*” [p. 4, -@wiernik2020]. They base their
claims largely on work done by Hunter and Schmidt [-@hunter2004], who
claim that “*Variation in reliability across studies causes variation in
the observed effect sizes above and beyond that produced by sampling
error.*” (p. 302). Overall, both references imply that differences in
score (un)reliability inevitably lead to an inflated heterogeneity in
ES.

As both sources Hunter and Schmidt [-@hunter2004] and Wiernik and Dahlke
[-@wiernik2020] will be referred to repeatedly, we abbreviate these
references as H&S and W&D respectively.

### Attenuation correction

If information concerning score reliability is available, it is possible
to correct the individual ES for its unreliability. This process is also
known as attenuation correction and, while not without its criticisms
[@muchinsky1996; @schmidt1996; @schennach2016], is a widespread practice
in Psychology (reference, alt.: repeatedly recommended + references).
@eq-dc describes a simple attenuation correction procedure.

$$d_c = \frac{d}{\sqrt{\hat{\rho}_{xx’}}}$$ {#eq-dc}

Here, $d_c$ describes the attenuation-corrected estimate of $d$, which
is corrected by dividing it by the estimate of score reliability
$\hat{\rho}_{xx’}$. As described in W&D or Lord & Novick [-@lord2008],
this is essentially the same, as if the corrected estimate of $d_c$ was
constructed using an estimate of true score standard deviation, as in
@eq-dct.

$$d_c = \frac{MD}{\hat{\sigma}_{T}}=\frac{MD}{\sqrt{\hat{\rho}_{xx’} \hat{\sigma}_X^2}}$$ {#eq-dct}

$\hat{\sigma}_T$ can be estimated by simply rearranging the terms from
@eq-rel. Thereby, an estimate of ES is generated, under the premise that
score reliability would have been perfect ($\rho_{xx'} = 1$).

The claims made in W&D imply that corrections of individual observed ESs
for their unreliability should lead to lower estimates of heterogeneity
in meta-analyses of corrected ES. If heterogeneity in score reliability
adds to ES heterogeneity, attenuation correction procedures should
eliminate that additional variation, as all corrected ES come with
identical (perfect) score reliability.

Re-analyses of the collaborative multi-site replication efforts have
uncovered differences in score reliability across replications of the
same phenomena [@shaw2020]. The claims made in W&D and H&S imply that
the ES heterogeneity uncovered in these projects therefore could
potentially be explained by these differences in score reliability.
Similarly, attempting to explain the substantial extent of ES
heterogeneity identified in psychological phenomena, both the original
authors as well as the multi-site replication authors claim that
differences in score reliability might be responsible (again, do we
actually know of any?).

Therefore, we assess the extent to which this intuition is supported by
empirical evidence in the following section. The re-analysis of archival
data serves to shed light on the following research question:

-   To what extent can heterogeneity in standardized effect sizes be
    reduced by means of an attenuation correction?

## Re-analysis of archival data

We have collected a large number of openly available multi-site direct
replications on psychological phenomena [@dripht2024]. The data-sets of
about 55 phenomena, largely stemming from the efforts of the Many Labs
studies, Registered Replication Reports or the Psychological Science
Accelerator, are made openly available in a standardised format at
\[<https://osf.io/7p8rk/>\], the DRIPHT repository. A documentation of
the standardised format, including a web-application to easily explore
the public data-sets, can be found in Fünderich et al.
[-@fuenderich_metapipex2024].

### Data

The ManyLabs projects were collaborative efforts to replicate several
psychological phenomena across different research sites, employing
identical protocols. From five published projects, the data for the
first three and the fifth Many Labs projects [@ManyLabs1; @ManyLabs2;
@ManyLabs3; @ManyLabs5] were harmonized and publicly stored in a
standardized format when the DRIPHT repository was set up [@dripht2024].
The Registered Replication Reports are similar collaborative efforts,
with the sole distinction that for each report a single phenomenon was
replicated several times. The Registered Reports 3-10 were added to the
DRIPHT repository, as they employed experimental designs with at least
two groups [@RRRHart; @RRRSripada; @RRRFinkel; @RRRStrack; @RRRRand;
@RRRDijksterhuis; @RRRSrull; @RRRMazar]. Lastly, some projects from the
Psychological Science Accelerator were added to the DRIPHT repository,
namely PSA006, PSACR001, PSACR002 and PSACR003 [@PSA006; @PSACR001;
@PSACR002; @PSACR003]. In contrast to Many Labs or Registered
Replication Projects, these collaborative efforts, while also
distributed across the globe, do not formally attempt to perform direct
replications. Instead, these projects focus on original research or
conceptual research done collaboratively in laboratories across the
globe. However, as each project used the same protocol and participants
are distributed across different countries, a data structure similar to
that of a multi-site replication study emerges. What makes all these
collaborative efforts valuable for this manuscript is the fact that all
phenomena collected in the DRIPHT-repository are making use of two-group
designs and are therefore easily assessed using standardised mean
differences.

While replication data on a decent number of psychological phenomena is
available in the DRIPHT-repository, selected phenomena need to fulfil
three conditions to demonstrate how incorporating score reliability
affects heterogeneity in ES: (1) As we focus on the use of ES $d$, the
phenomenon needs to be assessed using a group design consisting of at
least two groups. (2) Score reliability estimates can be derived, this
implies that the phenomenon needs to be measured using either several
indicators forming a single scale or by repeatedly measuring across
several timepoints. (3) Score reliability can only attenuate effect
sizes that are not zero in the first place. Therefore, we focus on
phenomena where meta-analytic mean ES can be statistically distinguished
from zero. This aligns with the observation made in (Renkewitz,
Fünderich & Beinhauer, in preparation) and Olsson-Collentine et al.
[-@olsson2020]. As null-effects displayed little to no heterogeneity,
there is no variability that might be reduced by incorporating
differences in score reliability. While the first criterion is fulfilled
by all phenomena found in the DRIPHT repository, the second and third
criteria still need evaluation.

As all phenomena in the collection fulfil the first condition, 50
phenomena have been catalogued where the effect is studied by comparing
the outcome across two separate groups. From the collection of 50
phenomena, 21 fulfil the second condition, so that estimates of score
reliability can be derived. For the 21 phenomena, the dependent variable
used to construct the effect size is measured by employing more than a
single indicator. The number of indicators per scale varied between 3
and 30. The majority of scales was made up of 3 to 6 items. At the same
time, however, this implies that the remaining 29 phenomena did not make
use of more than a single indicator to measure the dependent variable.
Therefore, for the majority of phenomena it is not possible to assess in
how far measurement quality is sufficiently high in terms of e.g.
internal consistency or low random error variance.

From the 21 phenomena that employed more than a single indicator, 19
made use of Likert-style items, where respondents would indicate their
agreement to some kind of statement. The scale-length varied from 3 to
10-points. The items measuring the remaining 2 phenomena were coded
dichotomously, as responses were right or wrong. While the majority of
designs make use of some form of control-treatment manipulation,
randomly assigning participants to a condition (19 phenomena), some
phenomena refer to the effect of pre-existing differences, e.g.
biological sex (2 phenomena). Subsequently, those phenomena need to be
identified where the meta-analytic mean effect size is statistically
significant different from zero. This will be done as a preliminary step
in this manuscript’s analysis procedure. Detailed information concerning
the different phenomena and how they were measured can be found at
\[osf-link\], while a brief summary can be found in @tbl-data.

To assess whether ES heterogeneity can be reduced by correcting for
score reliability, estimates of meta-analytic heterogeneity are compared
across two situations: first, heterogeneity of raw ES, as computed in
@eq-d is assessed. Secondly, ES are corrected for imperfect reliability.
Computing the heterogeneity of these corrected ES allows for an
assessment, whether the heterogeneity did indeed shrink compared to the
first situation, as predicted in W&D and H&S.

In order to derive estimates of meta-analytic heterogeneity, we make use
of a random-effects meta-analysis. Broken down, a random-effects
meta-analysis essentially pools the sampling uncertainty of the
different estimates that are to be aggregated. By removing this pooled
estimate from the observed variability in the effect size estimates, we
can generate an estimate of its heterogeneity - how strongly the effect
sizes vary beyond their pooled sampling uncertainty.

For the standardized mean difference $d$, we can compute the standard
error as an estimate of its sampling uncertainty. The standard error of
Cohen’s $d$ is computed as defined in @eq-SEd.

$$SE_d=\sqrt{\frac{n_1+n_2}{n_1n_2}+\frac{d^2}{2\left(n_1+n_2\right)}}$$ {#eq-SEd}

As the corrected estimate of Cohen’s $d$ is larger and requires an
additional unknown parameter, an estimate of score reliability, the
larger uncertainty in this parameter should be reflected in a larger
standard error. The standard error for $d_c$ is defined in @eq-SEdc.

$$SE_{d_c} = \sqrt{SE^2_d \frac{d_c}{d}}$$ {#eq-SEdc}

The estimates concerning the meta-analytic mean and heterogeneity of
these standardized effect sizes attempt to describe their population
distribution. Therefore, over the following paragraphs, whenever we
refer to estimates derived by means of random effects meta-analysis, we
will use the Greek letter $\delta_0$ to denote estimates of uncorrected
$d$'s population distribution and $\delta$ to denote estimates of
corrected $d_c$'s population distribution.

### Methods

For each replication of a phenomenon, Cohen’s $d$ is computed as an
estimate of raw ES, according to @eq-d, including its standard error
(@eq-SEd). Subsequently, a random-effects meta-analysis is performed
using metafor version 4.8-0 [@metafor]. To estimate meta-analytic mean
ES and heterogeneity we make use of the REML-estimator, as that
estimator tends to perform best across a range of conditions
[@hoenekopp2022]. To identify which phenomena pass criterion (3) we use
a Wald-type significance test, assessing for which phenomenon the
meta-analytic mean ES is significantly different from zero. For this
hypothesis test, as all other hypothesis tests in this manuscript, we
employ a one-tailed significance level of .05.

Additionally, estimates of score reliability are derived. While not
without criticism [e.g. @sijtsma2009], Cronbach’s Alpha is used to
estimate score reliability. It is often noted that tau-equivalence is an
unrealistic assumption to hold against real-world data, implying that
other estimates of score reliability, such as McDonald’s Omega
[@mcdonald1999], Guttman’s Lambda 2 or 4 [@guttman1945], or the Greatest
Lower Bound [@tenBerge2004] might be better suited. However, all scales
employed in this project are scored by computing the simple mean or sum
of responses across items for each individual. Computing the mean or sum
of several items implicitly assumes tau-equivalence, as each item
contributes equally to the individual’s test score. Therefore,
Cronbach’s Alpha is the better choice to estimate score reliability, as
it avoids a mismatch in assumptions between how the score is computed
and how score reliability is estimated. The individual estimates of
Cohen’s $d$ are corrected using estimates of Cronbach's Alpha, according
to @eq-dc, including the corrected standard errors (@eq-SEdc). Thereby,
estimates of ES are computed, which are corrected for imperfect
reliability. Again, a random-effects meta-analysis is run using metafor,
generating an estimate of heterogeneity in corrected ES. Different
indicators of heterogeneity are readily available, such as $I^2$, $H^2$
or the coefficient of variation $CV$[@borenstein2009; @higgins2002].
However, the descriptions found in H&S and W&D discuss the absolute
amount of heterogeneity in terms of variance ($\tau^2$) or standard
deviation ($\tau$). Therefore, we also make use of absolute
heterogeneity here.

Subsequently, in order to assess whether ES heterogeneity was indeed
reduced by the attenuation correction procedure, the estimates of
heterogeneity in uncorrected ES and corrected ES are compared. Using
Cochran’s *Q*-test [@cochran1954], we assess whether the estimates of
heterogeneity are statistically significantly different from 0. Only for
those projects where we have sufficient confidence that the individual
estimates of heterogeneity are larger than zero can we actually begin to
interpret whether there was any change. \[Short paragraph discussing
that a test for difference in heterogeneity would be more informative?
Maybe move it to the discussion section - hint at it here?\]

To identify whether a reduction in ES heterogeneity is indeed
accompanied by differences in score reliability, the estimates of score
reliability are assessed meta-analytically as well. To do so, a
Reliability Generalization Meta-Analysis is performed [@vacha1998;
@sanchez2013]. This entails that individual estimates of score
reliability are adequately transformed using Bonett’s transformation:

$$T_{\hat{\rho}} = ln(1 - \hat{\rho})$$ {#eq-Bonett}

The transformation in @eq-Bonett has variance-stabilising properties and
therefore allows for adequate inferences concerning heterogeneity in
score reliability [@bonett2002]. Cochran’s Q-test is used to identify
statistically significant heterogeneity in transformed score
reliability. Estimates derived from a Reliability Generalization
Meta-Analysis using these transformations can be back-transformed to the
original score reliability scale [@beinhauer2025]:

$$\mu_{\rho{xx'}} \approx 1 - exp(\mu_{T\left[\rho_{XX^\prime}\right]})-\frac{1}{2} exp(\mu_{T\left[\rho_{xx'}\right]})\tau_{T\left[\rho_{XX^\prime}\right]}^2$$ {#eq-rel_back_mu}

In @eq-rel_back_mu, $\mu_{\rho_{XX'}}$ is the back-transformed
meta-analytic mean score reliability - on the original coefficient's
scale between 0 and 1. Similarly, $\tau_{\mu_{\rho_{XX'}}}$ is the
back-transformed heterogeneity in score reliability:

$$\tau_{\rho_{xx'}}^2 \approx exp(\mu_{T\left[\rho_{xx'}\right]})2 \tau_{T\left[\rho_{xx'}\right]}^2+exp(\mu_{T\left[\rho_{xx'}\right]})^2\tau_{T\left[\rho_{xx'}\right]}^4$$ {#eq-rel_back_tau}

We executed these steps for each phenomenon. The statistical programming
language R, Version 4.4.2 [@RCore] is used for all data manipulation and
statistical analysis. Graphics are constructed using the R package
*ggplot2*, version 3.5.1 [@ggplot2].

### Results

[Equations @eq-d], [-@eq-SEd], [-@eq-dc], and [-@eq-SEdc] were used to
generate estimates of standardized ES, both uncorrected and corrected,
with their corresponding standard errors. Generally, this leads to
larger (absolute) ES and larger standard errors after the correction. As
an example from the 21 selected phenomena, the results from this
procedure on replications of Nosek’s study on sex differences in
preferences for the subject of Arts is displayed in the forest plot in
@fig-forest. Here, grey dots represent the uncorrected ES for each
sample, while black dots represent the corrected ES in each sample. The
bars surrounding these dots show the respective 95%-Cis.

```{r, echo = FALSE, fig.height = 6}
#| label: fig-forest 
#| fig-cap: Forest Plot ML1 - replications of Nosek's studz on sex differences in subject preferences (Arts)
forest_plot_rel(ES_rma_df[which(ES_rma_df$MASC == "Nosek_Explicit_Art"),], agg_L[[which(MASC_names == "Nosek_Explicit_Art")]])
```

In @fig-forest it becomes apparent that the reliability attenuation
procedure leads to an increase in the individual absolute effect size,
as the black dots are moved further from zero. Simultaneously, the
standard errors grew larger after the attenuation correction, which
leads to larger 95%-CIs. This is most easily observed for the
MTurk-sample, which had a rather large confidence interval to begin
with. Additionally, in the diamond at the bottom of the figure, we can
see that the meta-analytic mean ES is also larger after the attenuation
correction took place. @tbl-meanES contains estimates and test results
for all phenomena, both corrected and uncorrected.

#### Which phenomena pass criterion 3)?

@tbl-meanES demonstrates that what we observed in @fig-forest holds
across all 21 phenomena. All meta-analytic effects are larger in their
absolute size after applying an attenuation correction procedure.
Similarly, the uncertainty quantified in the estimate's standard error
is larger. Additionally, @tbl-meanES highlights which phenomena pass
criterion (3)—the effect size must be statistically significantly
distinguishable from zero at $p < .05$. 12 phenomena pass this
criterion. It may be remarked that the attenuation correction procedure,
in this data, appears to have no influence on whether a phenomenon
passes the criterion. The differences in p-value are rather small and
lead to no difference in conclusion, regarding a significance level of
5%.

```{r, echo = FALSE}
#| label: tbl-meanES
#| tbl-cap: Meta-analytic average  ES

gt::gt(read.csv(here("Tables/Mean_ES.csv")) %>% 
         arrange(MASC)) %>% 
  gt::cols_label(MASC = "MASC",
                 #mu_str_raw = gt::md("$\\hat{\\mu}_{\\delta_0}$ (SE)"),
                 mu_str_raw = gt::md("Mean $d$ (SE)"),
                 pval_raw = "p",
                 #mu_str_cor = gt::md("$\\hat{\\mu}_{\\delta}$ (SE)"),
                 mu_str_cor = gt::md("Mean $d_c$ (SE)"),
                 pval_cor = "p") %>% 
  gt::tab_options(table.width = gt::pct(100),
                  column_labels.font.weight = "bold") # %>% 
  # gt::cols_width(pval_raw ~ gt::px(40),
  #                pval_cor ~ gt::px(40))
```

*Note:* In this table *Mean* $d$ represents the meta-analytic estimate
of mean uncorrected $d$ and *Mean* $d_c$ represents the estimate of the
meta-analytic mean of corrected $d_c$.

#### Estimates of heterogeneity

@tbl-tauES contains the results from performing an RG-MA on the
Cronbach's Alpha coefficients of the scales used to measure the 12
remaining phenomena. Additionally, it contains estimates of
heterogeneity for both uncorrected ES ($\delta_0$) and corrected ES
($\delta$), which are central to assess the research question posed in
section *Attenuation Correction*. Most importantly, @tbl-tauES
demonstrates that for all but one of the phenomena, the attenuation
correction procedure barely has an impact on the extent of ES
heterogeneity. However, while the heterogeneity for the phenomenon from
RRR9 (Behaviour) seems to be reduced by about a third of its initial
value, the ES heterogeneity for this phenomenon is rather small to begin
with. Using a significance test, we can not distinguish this value
confidently from zero. This implies that for the majority of phenomena,
differences in score reliability can not be made responsible for
heterogeneity in standardized effect sizes.

```{r, echo = FALSE}
#| label: tbl-tauES
#| tbl-cap: Tests for ES Heterogeneity
#knitr::kable(x = read.csv(here("Tables/Heterogeneity_ES.csv")))

tab_hetES <- read.csv(here("Tables/Heterogeneity_ES.csv"))
tab_hetrel <- read.csv(here("Tables/Results_RMA_alpha.csv"))

tab_both <- data.frame(tab_hetES$MASC,
                       tab_hetrel[,c("mu_alpha_str", "tau_alpha", "QE_str", "QEp")],
                       tab_hetES[,c("tau_raw", "QE_raw_str", "QEp_raw", "tau_cor", "QE_cor_str", "QEp_cor")])

# knitr::kable(tab_both,
#              col.names = c("MASC", 
#                            "$\\hat{\\mu}_{\\rho_{XX'}}$ (95%-CI)", 
#                            "$\\hat{\\tau}_{\\rho_{XX'}}$", 
#                            "QE (df)", "p", 
#                            "$\\hat{\\tau}_{\\delta_0}$", 
#                            "QE (df)", "p",
#                            "$\\hat{\\tau}_{\\delta}$", 
#                            "QE (df)", "p"),
#              escape = FALSE)

gt::gt(tab_both  %>% 
  arrange(tab_hetES.MASC)) %>% 
  gt::cols_label(tab_hetES.MASC = "MASC",
                 mu_alpha_str = gt::md("$\\hat{\\mu}_{\\rho_{XX'}}$"),
                 tau_alpha = gt::md("$\\hat{\\tau}_{\\rho_{XX'}}$"),
                 QE_str = "QE (df)",
                 QEp = "p",
                 tau_raw = gt::md("$\\hat{\\tau}_{\\delta_0}$"),
                 QE_raw_str = "QE (df)",
                 QEp_raw = "p",
                 tau_raw = gt::md("$\\hat{\\tau}_{\\delta}$"),
                 QE_raw_str = "QE (df)",
                 QEp_raw = "p") %>% 
  gt::tab_options(table.width = gt::pct(100),
                  column_labels.font.weight = "bold") #%>% 
  # gt::cols_width(pval_raw ~ gt::px(40),
  #                pval_cor ~ gt::px(40))

# tab_one <- data.frame(tab_muES,
#                       tab_hetrel[,c("mu_alpha_str", "tau_alpha", "QE_str", "QEp")])
```

*Note:* $\hat{\mu}_{\rho_{XX'}}$ is the back-transformed meta-analytic
estimate of mean score reliability; $\tau_{\mu_{\rho_{XX'}}}$ is the
back-transformed estimate of heterogeneity in score reliability;
$\hat{\tau}_{\delta_0}$ represents the estimate of heterogeneity in
uncorrected $d$; $\hat{\tau}_{\delta}$ represents the estimate of the
heterogeneity in corrected $d_c$.

Secondly, in @tbl-tauES we find that for ten out of twelve phenomena, ES
heterogeneity grows as a result from the attenuation correction
procedure. However, it needs to be noted that these differences are small
and could not be distinguished from sampling uncertainty in a statistical
test. For this reason, we do not mean to interpret these differences, while 
we note that the direction of differences are clearly inconsistent with
expectations formulated in the literature discussed so far.
These results are incompatible with claims made in W&D and H&S.

Besides the direction of these differences, it is noteworthy that considering
differences in score reliability barely affects heterogeneity in standardized
effect sizes. [ausführen, oder kommt das noch einmal wieder?]

Concerning the
remaining two phenomena, for the phenomenon described in ML5 - Shnabel
(RPP), the reduction in heterogeneity is miniscule and the estimate of
heterogeneity is not statistically significant in the first place. Most
likely, the difference in ES heterogeneity estimates can be explained by
issues of estimation, as only eight replications are available.
Similarly, only for the phenomenon described in RRR9 (Behaviour), we
find a seemingly substantial reduction in heterogeneity due to the
attenuation correction procedure. The estimates of ES heterogeneity are
not statistically significant, however, which hampers our ability to
meaningfully interpret and compare these estimates. For the majority of
phenomena it appears that differences in score reliability do not
inflate heterogeneity in uncorrected effect sizes. Instead these
differences mask ES heterogeneity, that can be unveiled by accounting
for score reliability.

## Relationship between score reliability and effect size heterogeneity

In the re-analysis of archival data in the previous section we
demonstrated that, at least for the multi-site direct replication and
PSA-projects, score reliability does not hold any noteworthy value as a
potential moderator explaining ES heterogeneity. Additionally, we
demonstrated that claims formulated in W&D and H&S regarding how score
reliability heterogeneity increases ES heterogeneity are false. We found a large
number of phenomena where, contrary to popular claims, correcting for
score reliability increases ES heterogeneity, instead of decreasing it.

This implies that there is an error in the reasoning or intuition behind
claims made in H&S and W&D, as our observations can not be explained by
the current literature. Over the following paragraphs, we will briefly
explore the intuition that might have led to the previous claims.
Subsequently, we will use analytical arguments to clarify how score
reliability actually affects ES heterogeneity. Therein, we also identify
potential conditions under which an attenuation correction will actually
increase ES heterogeneity - and under which it is more likely to
decrease it.

### Intuitive relationship

It might help to think about linear models, when discussing the
intuitive idea that controlling for score
reliability should reduce ES heterogeneity. Heterogeneity describes
unexplained variation in a parameter's true distribution, that we model
using meta-analytical techniques [@borenstein2009; @higgins2002]. In
that way, the idea of heterogeneity is closely related to residual
variance, as we observe it in a simple linear model. Both describe
variation that could not be explained by the model specifications, be it a model that 
employs predictive variables like a simple linear model or a meta-analytic
model incorporating uncertainty in terms of standard errors.

The residual variance that remains in a simple linear model will be reduced
in all instances, where we use a predictor that we know shares 
variance with the predicted parameter. Generally, removing a source of variance
by controlling for it, we expect any unexplained variation to shrink.
Similarly, we know that differences in score reliability, by definition,
are related to differences in standardized effect sizes. Therefore,
intuitively, it seems sensible to expect that controlling for
differences in score reliability should reduce the unexplained variance
in standardized effect sizes, as described in H&S and W&D.

Hunter & Schmidt have additionally derived an equation which
demonstrates how heterogeneity in score reliability affects
heterogeneity in ESs [-@hunter2004, p. XX]. @eq-HS implies that
heterogeneity in $\delta_0$ is essentially a function in which
heterogeneity and mean value of $\delta$ are weighted by mean score
reliability $\rho_{xx‘}$ and its heterogeneity.

$$\tau^2_{\delta_0} = [\mu_{\rho_{xx‘}}]^2 \tau^2_{\delta} + [\mu_{\delta}]^2 \tau^2_{\rho_{xx‘}}$$ {#eq-HS}

We adjusted @eq-HS using the notation used throughout this text. Here,
$\delta_0$, carrying a 0 in the index, refers to the uncorrected ES,
undistorted by sampling error but not corrected for measuring error. Without an 
index, $\delta$, on the other hand, refers to the corrected ES,
undistorted by sampling error and measurement error. Lastly,
$\rho_{XX'}$ refers to score reliability. Hunter & Schmidt
[-@hunter2004] claim that @eq-HS demonstrates that differences in score
reliability inflate ES heterogeneity, which means that an attenuation
correction procedure would have to reduce ES heterogeneity. However, it
can be easily verified that @eq-HS does not allow for such claims.
Depending on both the mean value $\mu_{\rho_{xx‘}}$ and the
heterogeneity $\tau^2_{\rho_{xx‘}}$ in score reliability, ES
heterogeneity might be increased or decreased by the differences in
score reliability.

Most likely, Hunter and Schmidt did not fully explore the implications
of @eq-HS, but followed the general intuition we lined out–"If a source
of variation is controlled for, the amount of unexplained variance
should be reduced." Wiernik and Dahlke [-@wiernik2020] refer to the work
by Hunter & Schmidt [-@wiernik2020], likely following their intuition.

In the re-analysis of archival data, we have demonstrated that this
intuition does not hold for the relationship of ES heterogeneity and
score reliability. Instead, we propose an alternative way that helps us
discern, under which conditions we can expect an attenuation correction
procedure to lead to an increase in ES heterogeneity.

### Describing ES as a random ratio variable

Realising that an ES, as defined in @eq-d, is actually a ratio, allows
for a different analytical description of how differences in score
reliability can affect ES heterogeneity. If the phenomenon is
heterogeneous and score reliability varies across replications, it seems
sensible to assume that both numerator and denominator (MD and
$\sigma_X$) from @eq-d vary across replications. In that case, the ES
can be described as a random variable stemming from a ratio
distribution. A ratio distribution is a probability distribution
constructed by dividing one random variable by a second random variable:
$Z = \frac{X}{Y}$. First order taylor approximation may be used to
generate estimates of mean and variance of the ratio variable
[@vanKempen1999]. Taylor approximations typically require no specific
assumptions regarding the distribution form, but that the variable (here
a ratio) is a) differentiable, b) has existing moments and that c) its
higher-order moments are negligible (e.g. limited variability and
monotonicity) (Asymptotic statistics, A.W. van der Vaart - find book online). Here we use the
Taylor-estimator of the ratio’s variance to demonstrate how differences
in score reliability may affect heterogeneity in ES.

$$\tau^2[Z] \approx \frac{\tau^2[X]}{\mu[Y]^2} - \frac{2\mu[X]}{\mu[Y]^3} cov[X,Y] + \frac{\mu[X]^2}{\mu[Y]^4} \tau^2[Y]$$ {#eq-XYZ_cov}

Assuming that the random variables are uncorrelated, simplifies the
equation to

$$\tau^2[Z] \approx \frac{\tau^2[X]}{\mu[Y]^2} + \frac{\mu[X]^2}{\mu[Y]^4} \tau^2[Y]$$ {#eq-XYZ}

Substituting for ES, unstandardized mean difference MD and pooled
standard deviation $\sigma_X$ leads to

$$\tau^2[\delta_0] \approx \frac{\tau^2[MD]}{\mu[\sigma_x]^2} + \frac{\mu[MD]^2}{\mu[\sigma_x]^4} \tau^2[\sigma_x]$$ {#eq-ratio_d0}

As stated in @eq-dct, correcting ES for score reliability can be
described through a correction of the pooled standard deviation:
$\sigma_T = \sqrt{\rho_{xx’} \sigma^2_x}$. Since $\rho_{xx'}$ is always
between zero and one, in the case of imperfect score reliability,
$\sigma_T$ will always be smaller than total $\sigma_x$.

In @eq-rel we see that total standard deviation $\sigma_X$ is
constructed only from error score variance
$\sigma_E^2$ and true score variance $\sigma_T^2$. Measurement
precision should generally only
affect the error score variance. Therefore, if the heterogeneity in
$\sigma_X$ was introduced purely by differences in measurement
precision, an attenuation correction procedure can remove that
heterogeneity. Alternatively, the underlying latent variable may not be
distributed identically across replications. In that case, true score
variance $\sigma^2_T$ would vary across replications and induce
heterogeneity in $\sigma_X$. This heterogeneity would lead to
differences in standardized differences ES as well, but it can not be
removed by means of an attenuation correction procedure.

In that case, some heterogeneity in $\sigma_T$ would remain, albeit
heterogeneity found in $\sigma_x$ was reduced. @eq-ratio_d0 essentially
demonstrates how mean and heterogeneity of MD, paired with mean and
heterogeneity of total score standard deviation $\sigma_x$ can be used
to estimate heterogeneity in raw ES $\delta_0$. Similarly, we can adjust
@eq-ratio_d0 to describe how, instead of total score standard deviation,
mean and variance in true score standard deviation $\sigma_T$ affect
heterogeneity in corrected ES $\delta$.

$$\tau^2[\delta] \approx \frac{\tau^2[MD]}{\mu[\sigma_T]^2} + \frac{\mu[MD]^2}{\mu[\sigma_T]^4} \tau^2[\sigma_T]$$ {#eq-ratio_d}

The differences between the two [Equations @eq-ratio_d0] and
[-@eq-ratio_d] demonstrates how ES heterogeneity is affected by an
attenuation correction procedure. According to the quotes in H&S and
W&D, @eq-ratio_d0 should lead to a larger value of
$\tau^2\left[\delta_0\right]$, compared to the estimate of
$\tau^2\left[\delta\right]$ from @eq-ratio_d. However, as demonstrated
in the re-analysis of archival data, we can already expect that this can
not be an implication of these Equations.

Instead, what is guaranteed is that the expected value of $\sigma_T$ is
smaller than or equal to the expected value of $\sigma_X$. Similarly, the
heterogeneity in $\sigma_T$ is guaranteed to be smaller than or equal to the
uncorrected heterogeneity in $\sigma_X$. However, since the expected
value of either standard deviation is placed in the denominator of the
function and the variance of either is placed in the numerator of the
function, @eq-HS is not sufficient to explain how heterogeneity in ES
changes due to the attenuation correction.

In this section, we attempt to understand under which circumstances the
heterogeneity in ES grows larger as the result of an attenuation
correction procedure, contrary to claims made in W&D and H&S. To do so,
we propose the following inequality, involving [Equations @eq-ratio_d0]
and [-@eq-ratio_d].

$$\tau^2_{\delta} > \tau^2_{\delta_0} \quad  \quad \quad \quad \quad  \frac{\tau_{MD}^2}{\mu_{\sigma_T}^2} + \frac{\mu_{MD}^2}{\mu_{\sigma_T}^4} \tau^2_{\sigma_T} > \frac{\tau_{MD}^2}{\mu_{\sigma_x}^2} + \frac{\mu_{MD}^2}{\mu_{\sigma_x}^4} \tau^2_{\sigma_x}$$ {#eq-ineq_init}

As long as the inequality defined in @eq-ineq_init holds true, an
attenuation correction procedure would lead to larger estimates of ES
heterogeneity. Conditions under which @eq-ineq_init does not hold are
conditions under which the claims made in W&D and H&S are met.
Unfortunately, the inequality in this form does not
convey sufficient information about the conditions under which it can
potentially hold. In Appendix A, we rearranged @eq-ineq_init, so we end
up with a form that we can use to actually derive conditions under which
ES heterogeneity can be inflated or deflated by differences in score
reliability.

To do so, we had to introduce two additional metrics.

$$R_1 = \frac{\mu_{\sigma^2_T}}{\mu_{\sigma^2_X}}$$ {#eq-R1}

$R_1$ describes the relative size of mean true score variance, compared
to the mean observed score variance. This informs us, how much of the
mean total score variance can be attributed to mean true score variance.
Therefore, this metric is close to the average score reliability. Even
though formally $R_1$ and the meta-analytic mean score reliability
$\mu_{\rho_{xx'}}$ do not share not the same value, in practice we
observe little to no differences between the two parameters. [add to appendix]. Large
values indicate that correcting the individual score variances (or
standard deviations for that matter) using the attenuation correction
procedure should lead to very little change in ES, as most of the
variation observed were due to genuine differences in true scores. Small
values indicate the opposite, the observed score variance was due to a
larger amount of random error score variance, leading to a stronger
correction. The metric $R_2$ on the other hand describes in how far the attenuation
correction procedure has successfully reduced heterogeneity in score
variance $\tau^2_{\sigma^2_X}$.

$$R_2 = \frac{\tau^2_{\sigma^2_T}}{\tau^2_{\sigma^2_X}}$$ {#eq-R2}

A value of 1 indicates that the
heterogeneity in true score variance $\sigma^2_T$ is exactly as
large as the heterogeneity initially observed. Smaller values indicate
how much heterogeneity is "left", after applying an attenuation
correction, relative to the initial heterogeneity. This means that a
value of .7 indicates that about 70% of heterogeneity in score variance
remains, even after correcting for differences in score reliability. As
only 30% of score variance heterogeneity could be removed, this would
imply that differences in error score variance $\sigma^2_E$ were
responsible for less than a third of the heterogeneity in score
variances found. On the contrary, more than two thirds of heterogeneity
could be attributed to actual differences in how the underlying true
scores are distributed across samples.

Concerning the re-arranged inequality from Appendix A, we identify one
sufficient but not necessary condition or rule described in
@eq-ineq_short.

$$R_1^3 \leq R_2$$ {#eq-ineq_short}

With sufficient but not necessary we mean that, as long as
@eq-ineq_short holds, an attenuation correction will inevitably increase
ES heterogeneity. However, if the inequality does not hold, it is not
guaranteed that the attenuation correction can decrease ES
heterogeneity. Whether ES heterogeneity can be decreased depends on
additional parameters, that we left out of this short form, but can be found,
accompanied by a more detailed description, in @eq-ineq_fin, Appendix A.

As long as $R_1^3$ is smaller or equal compared to $R_2$, @eq-ineq_init
is bound to hold true. This means that as long as the average score
reliability, to the power of 3, does not exceed the relative remaining
variance heterogeneity after the attenuation correction, @eq-ineq_init
is not violated. For example, assuming average score reliability is
about .8, the heterogeneity in $\sigma^2_T$ needs to make up less than
51.2% ($R_2 \geq .8^3 = .512$) of the heterogeneity in total score
variance $\sigma^2_X$. At $R_1 = .8$, as long as $R_2 \geq .512$, the
claims made in W&D and H&S can not hold.

$R_1$ essentially describes the relative size of mean score variance
after the attenuation correction, and $R_2$ essentially describes the
relative size of score variance heterogeneity after the attenuation
correction. Therefore, @eq-ineq_short states the following regarding the
change in mean value and heterogeneity due to the correction: ES
heterogeneity will always be increased, as long as the remaining proportion of
heterogeneity in score variance is not substantially smaller than the
remaining proportion of its mean value.

As long as there are substantial differences in the true score variance
across administrations, it is unlikely that an attenuation correction
can actually produce such a low $R_2$. Generally speaking, $R_1$ largely
depends on the relative measurement precision. If measurements were made
more precisely, average score reliability and $R_1$ would be higher. On
the other hand, $R_2$ depends on how different the measurement precision
was at the individual administrations, but also how different the
individual administrations sites are regarding their true score
variance. If the administration sites carry large differences in true
score variance $R_2$ is unlikely to be substantially smaller than $R_1$. [Na ja, eine Reliabilität von .9 bedeutet erstmal, dass es 9mal so viel wahre Varianz wie Fehlervarianz gibt. In diesem Fall müsste R2 unter 0.73 liegen. Das impliziert in der Heterogenität ein Verhältnis von ungefähr 3 : 1. Die 9mal größere wahre Varianz müsste also für weniger als ¾ der Heterogenität verantwortlich sein. In Deinem Textbeispiel (Reliabilität von .8) müsste die 4mal größere wahre Varianz für weniger als die Hälfte der Heterogenität verantwortlich sein. – Das erscheint mir beides per se ziemlich unplausibel… Also auch dann, wenn die „administration sites do NOT carry large differences in true score variance”. – In anderen Worten: Die Frage ist, ob die absolute Größe der Varianz überhaupt irgendeine Rolle spielt oder ob alles lediglich an den Verhältnissen hängt, die durch die Reliabilität vorgegeben sind. Ich habe dazu aber auch keine eindeutige Antwort… Wenn ich überzeugt wäre, dass sich *homogene* true score variances durch planvolle, kontrollierte Stichprobenziehung herstellen lassen, wäre das ein Gegenargument.]

It is important to reiterate that the condition in @eq-ineq_short only
describes under which conditions the inequality is guaranteed to hold,
no matter what value other parameters like the mean value of $MD$ or its
heterogeneity take on. This means that, if the condition in
@eq-ineq_short is violated, it is by no means guaranteed that an
attenuation correction procedure would actually reduce heterogeneity in
ES. Thereby, inference made from @eq-ineq_short is asymmetrical. While
we know, if the inequality holds, heterogeneity in corrected ES will be
larger than heterogeneity in uncorrected ES, we can't be certain that
the opposite is true if the inequality does not hold.

### Re-analysis of archival data

@tbl-tauES showed that for all phenomena, for which we identified
statistically significant ES heterogeneity, the heterogeneity was even
larger after the attenuation correction procedure. Here, we will
re-analyse the data, estimating $R_1$ and $R_2$ for each phenomenon
respectively. Therein, we not only show that our claims derived from
analytical arguments hold up in empirical assessment, but also
demonstrate how the metrics $R_1$ and $R_2$ help understand how
score reliability affected ES heterogeneity in practice.

#### Methods {#sec-methods2}

Estimation of $R_1$ and $R_2$ requires estimates of meta-analytic mean
and heterogeneity of both the total score variance $\sigma^2_X$ and the
true score variance $\sigma^2_T$. Elsewhere, we proposed the method
Boot-Err [@beinhauer2025] as an alternative method to test for and
estimate the differences in error score variance $\sigma^2_E$. Among
other things, in Beinhauer et al. [-@beinhauer2025] we deliver detailed
instructions on how estimates for mean and heterogeneity of total score
variance, and the score variance components can be derived. This means
that Boot-Err can be used to generate all estimates required for
computation of the metrics $R_1$ and $R_2$.

As Boot-Err makes use of
meta-analytic techniques and bootstrap sampling, we perform analyses
using the packages *metafor*, version 4.8-0 [@metafor], and *boot*,
version 1.3-31 [@boot], in the statistical programming language *R*,
version 4.4.2 [@RCore].

#### Results

@tbl-R1R2 reports the meta-analytic estimates of observed and true score
variances, including their heterogeneity and metrics $R_1$ and $R_2$
across all 12 phenomena. We see that for ten out of 12
phenomena, the $R_2$ metric is larger than the $R_1$ metric.
Additionally, we observe that across eight phenomena, $R_2$ is larger
than .96. This implies that of the heterogeneity observed in score
variances, 97% or more could not be explained by differences in random
error introduced by the measuring instrument, but by genuine differences
in the underlying distribution.

For the two exceptions in @tbl-tauES, those phenomena where the attenuation correction
procedure actually led to a decrease in ES heterogeneity (RRR9
(Behaviour) and ML5 - Shnabel (RPP)), we find that $R_2$ is larger than $R_1$.
This clearly contradicts our interpretation of @eq-ineq_int. As long as $R_2$ is
larger than $R_1^3$, we expect a correction procedure to increase effect 
size heterogeneity. However, we recommend to be cautious when interpreting both the change and
the direction of said change for these two projects. 
The significance test for the ES
heterogeneity was not significant, both before and after the correction.
Since either the heterogeneity was so small or the sample
size so low that the heterogeneity could not be statistically
distinguished from zero, most likely, the reduction in
heterogeneity is better explained by imprecise estimation than an actual
reduction of heterogeneity.

Lastly, for two out of the 12 phenomena, ML5 -
Albarracin and ML5 - Shnabel (Rev), we find
$\tau_{\sigma^2_X}$ of 0. Therefore, the $R_2$ metric could not be
computed, as the denominator would have been zero. However, we also did
not identify statistically significant heterogeneity in score reliability
across replications, nor ES heterogeneity. The phenomena were assessed with a 
particularly low number of
replications (9 and 8 respectively). Estimation precision for
heterogeneity in variance components, even more so than for ES, largely
depends on the sample size. Two alternative explanations for this observation
are available: either, there is no heterogeneity in score variances across 
replications. In that case, there is also no heterogeneity that could be
broken down into true and error score variance heterogeneity. Alternatively,
the low sample size led to
implausible estimates in both true and total score variance
heterogeneity and subsequently to an implausible estimate of $R_2$.
Meta-analytic estimates, especially those concerning score variance
components such as $R_2$, should generally be interpreted with caution,
if the sample size was particularly low.

Generally, @tbl-R1R2 implies that for none of the measurements concerning these
phenomena, the heterogeneity in observed score variances could be
substantially explained by differences in measurement quality. For all
phenomena, $R_2$ was larger than $R_1$, implying that an attenuation
correction procedure could not reduce ES heterogeneity in any of these
cases. The results in @tbl-tauES agree with this conclusion, as no
phenomena could be identified where statistically significant ES
heterogeneity was accompanied by a reduction in differences due to the
correction.

```{r, echo = FALSE}
#| label: tbl-R1R2
#| tbl-cap: Re-analysis concerning metrics $R_1$ and $R_2$
# knitr::kable(read.csv(here("Tables/Variances_analysis.csv")),
#              col.names = c("MASC", 
#                            "$\\hat{\\mu}_{\\sigma^2_X}$", 
#                            "$\\hat{\\mu}_{\\sigma^2_T}$", 
#                            "$\\hat{\\tau}_{\\sigma^2_X}$", 
#                            "$\\hat{\\tau}_{\\sigma^2_T}$", 
#                            "$R_1$", 
#                            "$R_2$"))

gt::gt(read.csv(here("Tables/Variances_analysis.csv")) %>% 
         arrange(MASC)) %>% 
  gt::cols_label(MASC = "MASC",
                 mu_X = gt::md("$\\hat{\\mu}_{\\sigma^2_X}$"),
                 mu_T = gt::md("$\\hat{\\mu}_{\\sigma^2_T}$"),
                 tau_X = gt::md("$\\hat{\\tau}_{\\sigma^2_X}$"),
                 tau_T = gt::md("$\\hat{\\tau}_{\\sigma^2_T}$"),
                 R1 = gt::md("R_1"),
                 R2 = gt::md("R_2")) %>% 
  gt::tab_options(table.width = gt::pct(100),
                  column_labels.font.weight = "bold") 
```

## Discussion

We have examined meta-analytic heterogeneity in standardized mean
differences. Since score reliability is often discussed as a source for
such heterogeneity, we estimated the heterogeneity before and after
correcting for differences in reliability coefficients using an
attenuation correction procedure. While we initially observed heterogeneity
in score reliability across almost all measuring scales, the extent of this
heterogeneity was very small across all scales. This initial finding already
implies that differences in score reliability can only have a tiny impact
on ES heterogeneity, if any.

This implication fits our subsequent finding, as we found that a) in the majority of
cases, the change in ES heterogeneity is small to non-existent. Across the 
data-sets we got to examine, differences in
score reliability do not serve as an appropriate explanation for ES
heterogeneity. We also found that b) in the majority of cases, the
direction of change in ES heterogeneity does not align with assumptions
formulated in W&D and H&S. In almost all cases (every single case, if we
restrict ourselves to phenomena where we found statistically significant
differences in ES), the attenuation correction led to increased
estimates of ES heterogeneity. This means that the claims found in W&D
and H&S, implying that an attenuation correction will reduce ES
heterogeneity, stem from a general misunderstanding of how score
reliability can affect ESs.

Using analytical arguments, we derived a broad rule, outlining under
which conditions we can expect an attenuation correction to increase ES
heterogeneity. In general, score reliability affects ESs by inflating
the score variance, as the (square root of the) score variance is used
to standardize the mean difference. The conditions we found depend on
how both the mean value and the heterogeneity of score variance are
reduced by correcting for imperfect score reliability. We introduced two
novel metrics: $R_1$, describing the fraction of mean true score
variance in the mean total score variance, therefore mirroring the
average score reliability coefficient, and $R_2$, describing the
fraction of true score variance heterogeneity in the total score
variance heterogeneity. We find that, as long as $R_1^3 \leq R_2$, an
attenuation correction will increase ES heterogeneity.

Since $R_1$ is carrying the exponent of 3, it implies that the reduction
in mean score variance is affecting the ES heterogeneity more strongly
than the reduction in score variance heterogeneity. This is best
described through an example: We know what an acceptable average score
reliability looks like in the area of Psychology. Researchers largely
target a score reliability of at least .7, preferably .8. An average
score reliability of .7 implies that 70% of the variance observed in
scores can actually be attributed to true underlying differences. From
our derived rule, we know that, as long as $R_2$ is larger or equal to
$.7^3 = .343$, the attenuation correction procedure can only increase ES
heterogeneity. If we want to explain ES heterogeneity through score
reliability, even though only 30% of the variance observed in score can
be attributed to random measurement error, at least 65.7%, almost two
thirds, of the heterogeneity in score variance would need to be
attributable to differences in the error score variance. In other words,
at most 34.3% of heterogeneity in score variances may be due to actual
differences in how the latent variable is distributed (in terms of
spread) across the administration sites. This disparity grows even larger, 
the higher the average score reliability is.

In the 
majority of research studies done in the social and behavioural sciences
[@hultsch2002; @bornstein2013; @ashraf2017; @sears1986; @zhao2021], no 
proper sampling techniques beyond convenience sampling are
employed. Most samples re-analysed throughout this text came to be
through convenience sampling, where students self-select into studies.
Without a valid sampling scheme, we essentially have no basis to
assume anything about the true score variance [@hultsch2002]. Most
importantly, however, we have no reason to assume that the true score
variance could potentially be homogeneous or stable across
administration sites. Instead, it is very likely that the convenience
sampling procedures produce samples that vary extremely strongly in how
they are put together, leading to large degrees of true score variance
heterogeneity. At the same time, score reliability is a low-hanging
fruit to optimise as most researchers are aware and trained in assessing
score reliability [@hussey2025].

The combination of these aspects can explain what we observed throughout
this text: Reliability coefficients are in an acceptable range, leading
to a small, but constant reduction in mean score variance if we apply an
attenuation correction. At the same time, there is substantial variation
in true score variance, implying that most of the score variance
heterogeneity can not be attributed to differences in error score
variance. Since only differences in error score variance can be
corrected for, the attenuation correction has a very small effect on
heterogeneity estimates in score variance. Therefore, an attenuation 
correction is likely to have very little
impact on ES heterogeneity. If anything, it most likely increases it.
This also means that, in such settings, score reliability does not serve
as a moderator explaining ES heterogeneity.

Our results fit in well with recent work by Olsson-Collentine et al.
[@olsson2023]. In a large simulation scheme, they demonstrate that
differences in score reliability across administrations typically
deflate heterogeneity in uncorrected correlations. In such cases, an
attenuation correction would increase correlation coefficient
heterogeneity, as we observe for standardized mean differences. Only as
the true heterogeneity in correlations grows larger would the correction
procedure reduce heterogeneity as implied in W&D. While correlations and
standardized effect sizes are not identical, the way score
(un)reliability affects these parameters is highly similar.

### Limitations & constraints on generalizability

The empirical arguments presented are based on a rather small number of
non-representative data-sets. Based on the combination of analytical and
empirical arguments, we are convinced that differences in score
reliability across replications in the current state of Psychology and
the behavioural sciences, do not serve as appropriate explanations of ES
heterogeneity. However, whether these results actually generalize beyond
these data-sets remains to be seen. Unfortunately, the behavioural
sciences are in dire need of open-data that resemble multi-site
replications. As the majority of results discussed over the last years
(in e.g. ManyLabs or Registered Replication Reports) employ
single-indicator scales as dependent measures, score reliability can not
be easily estimated in order to replicate our analyses.

The analytical arguments, detailed in *Appendix A*, leading to
@eq-ineq_short, rest on the use of the delta-method. @eq-ineq_init
contains parameters of the score standard deviations’ distributions,
while the $R_1$ and $R_2$ metrics were designed to discuss score
variances. The delta-method rests on a number of assumptions, at least
one of which is likely violated in our analysis: the variance of the
variable to be transformed needs to be small [@cramer1999]. This entire
procedure discusses the influence of variance in standard deviations on
ES heterogeneity, introduced by differences in score reliability.
Therefore, assuming that this assumption holds seems naive. In the
*Appendix B*, we explore the archival data, to understand in how far
employing the delta-method on meta-analyses of standard deviations lead
to biased results than the meta-analyses on score variances we performed
using the Boot-Err technique. The differences identified are rather
miniscule for the majority of phenomena, at the third decimal. Only for
the phenomenon in PSACR002 (Behaviour) we find larger differences -
therefore, conclusions for this phenomenon concerning the relation
between $R_1$ and $R_2$ need to be drawn with caution. Generally, as the
majority of phenomena did not display any bias in this assessment and
since the delta-method is not actually used to generate estimates, we
believe its use to be defensible nonetheless.

Similarly, all equations following @eq-XYZ_cov rest on the assumption
that mean difference ($MD$) and true or total score variance
($\sigma^2_T$ and $\sigma^2_X$) are independent. While this assumption
is in line with assumptions underlying CTT, it is by no means guaranteed
that it holds in these data-sets. If this assumption was violated,
whether the relationship was positive or negative, indicated by a
positive or negative covariance in @eq-XYZ_cov, would most likely have
an impact on how strongly ES heterogeneity is affected by the
attenuation correction procedure, possibly even the direction of how it
is affected. However, currently we have no reason to assume that there
is a systematic relationship between mean difference and standard
deviation. As this relationship would also affect the distribution of ES
like Cohen’s d itself, strong violations of this assumption would go far
beyond invalidating the claims we derived in @eq-ineq_short. Most
likely, assumptions of the meta-analytic tests concerning the
meta-analytic mean size ES and the presence of ES heterogeneity would be
violated, making any research on change in ES heterogeneity due to
attenuation correction obsolete.

## Conclusions

Heterogeneity in ES even in direct replications, where experimental
factors are held as constant as possible, is substantial
(Renkewitz, Fünderich & Beinhauer, in preparation). However, we
have demonstrated that in several scenarios, differences in score
reliability did not serve as appropriate moderators or explanations of
ES heterogeneity. Across most phenomena, we failed to reduce ES
heterogeneity. If anything, we even increased ES heterogeneity, albeit
to a very small degree. Generally, concerning the multi-site replication
projects or collaborative research projects, accounting for differences
in score reliability did not help us gain a better understanding of the
phenomena.

### Where to go from here

This manuscript has demonstrated that, even though differences in
measurement precision are present across most phenomena we could re-analyse, 
these are strongly outweighed by the differences in true score variance.
However, what a true score variance actually means, strongly depends on 
how that number came to be. A sample that is very diverse in its composition
likely comes with a large true score variance. On the other hand, a sample
that is highly homogeneous in its composition comes with a smaller true score
variance. Importantly, the same is not true for random measurement error. 
Error variance should always be the same size, if the errors are truly random,
no matter how the sample is put together. Unfortunately, 
most phenomena we re-analysed here came to be through means of convenience sampling. 
Therefore, it is not clear what the large differences in true score variance 
actually imply - are they expressions of
genuine differences between the populations, or just expressions of different
self-selection procedures across the samples? As the variance observed in
a convenience sample is not representative to some *specificied* population, 
we simply can not know.

Issues of generalizability
due to the lack of sampling techniques have been, and are still, repeatedly
called out for the same reasons. Our findings therein reiterate these calls. 
We believe that we should
focus our data collection efforts on the generation of samples, where we
know that the variance in a sample is actually meaningful. This is most
easily achieved by proper sampling techniques, which ensure that the
sample is representative for some specified population. In such cases, it might be
that the true score variance varies less strongly than what we currently
observe. And should the heterogeneity in true score variance remain as
large, than we know that this is crucial information from which we
can draw meaningful inferences. We could learn more about
standardized effect sizes and why our phenomena vary across settings, 
populations or continents. We strongly believe that the valuable resources 
invested into
research output could be put towards more informative use than what seems
to be the status quo by ensuring that the variances observed in different
samples are representative for specified populations.

The work done by Hunter and Schmidt [-@hunter2004] was crucial in
informing and guiding methodology development in the field of
meta-analyses. Similarly, Wiernik and Dahlke [-@wiernik2020] raise a
number of important points that go far beyond attenuation correction and
have been neglected in the application of meta-analytic research over
the last decades. We generally agree with their ideas that
underappreciated (differences in un-)reliability can introduce
substantial biases in meta-analytic estimates and tests in Psychology.
However, based on our findings, we believe that the issues created by
ignoring random sampling are much more dire and warrant immediate
attention. Only once we can be sure that the variance in true scores or
latent factors is actually meaningful for some population, we can turn
towards imprecise measurements, expecting correction procedures to
explain unwarranted heterogeneity. In the current state of Psychology,
meta-analysts should not expect that correcting ES for score reliability
will actually explain heterogeneity identified.

```{r, echo = FALSE}
#| label: tbl-data
#| tbl-cap: Brief summary of phenomena, designs and measures employed

tab <- read.csv("Tables/data_summary.csv", fileEncoding = "UTF-8", sep = ";") %>% 
  arrange(MASC)

gt::gt(tab)
```

## References

::: {#refs}
:::

## Appendix

### A - Rearranging inequality

Both metrics $R_1$ and $R_2$ make use of parameters of the distribution
of true and total score variance, not standard deviations. This choice
was made to facilitate the role of score reliability in the subsequent
discussion, as score reliability itself is defined by variances, not
standard deviations (see @eq-rel).

Using the metrics $R_1$ and $R_2$, defined in [Equations @eq-R1] and
[-@eq-R2], we attempt to understand under which circumstances the
inequality defined in @eq-ineq_init holds true. However, while the
metrics make use of mean and heterogeneity of the score variances,
@eq-ineq_init makes use of mean and heterogeneity of the standard
deviations. In order to incorporate the metrics, it is necessary to
reparameterise the equation. Using the delta method to approximate mean
value and heterogeneity of variance from those of the standard
deviations, we know that (ref):
$$\mu^2_{\sigma_X} \approx \mu_{\sigma^2_X} \quad \quad  and \quad \quad 
\tau^2_{\sigma_X} \approx \frac{\tau^2_{\sigma_X^2}}{4\mu_{\sigma^2_X}}$$ {#eq-delta_method}

While @eq-delta_method only explicitly contains parameters for the
observed score variance/standard deviation, the same can be done using
the true score variance/standard deviation. Using the approximates
defined in @eq-delta_method for observed and true score variance, we
arrive at a new inequality in @eq-ineq_CV.

$$\frac{\tau_{MD}^2}{\mu_{\sigma^2_t}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_t^2}} \frac{\tau^2_{\sigma_t^2}}{4\mu_{\sigma^2_t}} > \frac{\tau_{MD}^2}{\mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} \frac{\tau^2_{\sigma_x^2}}{4\mu_{\sigma^2_x}}$$ {#eq-ineq_CV}

@eq-ineq_CV describes that the heterogeneity in corrected ES is larger
than the heterogeneity in uncorrected ES, using parameters of the
distributions of mean difference MD, true score variance $\sigma^2_T$
and observed score variance $\sigma^2_X$. By introducing the metrics
$R_1$ and $R_2$ to this inequality, we can begin to disentangle which
conditions need to be fulfilled for the inequality to hold. Rearranging
the terms in [Equations @eq-R1] and [-@eq-R2], and subsequently entering
these terms into @eq-ineq_CV leads to the following
$$\frac{\tau_{MD}^2}{R_1 \mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{R_1^2 \mu^2_{\sigma^2_x}} \frac{R_2 \tau^2_{\sigma_x^2}}{4 R_1 \mu_{\sigma^2_x}} > \frac{\tau_{MD}^2}{\mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} \frac{\tau^2_{\sigma_x^2}}{4\mu_{\sigma^2_x}}$$ {#eq-ineq_R}

@eq-ineq_R, again, describes the inequality that heterogeneity in
corrected ES is larger than the heterogeneity in uncorrected ES,
incorporating the metrics $R_1$ and $R_2$. This equation alone is not
sufficient to identify the relevant circumstances required for the
inequality to hold. In a series of steps, we will rearrange @eq-ineq_R
to facilitate interpretation.

Firstly, for @eq-ineq_rearr1 we pulled all terms onto the left side of
the inequality.

$$\frac{\tau_{MD}^2}{R_1 \mu_{\sigma^2_x}} - \frac{\tau_{MD}^2}{\mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{R_1^2 \mu^2_{\sigma^2_x}} \frac{R_2 \tau^2_{\sigma_x^2}}{4 R_1 \mu_{\sigma^2_x}} - \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} \frac{\tau^2_{\sigma_x^2}}{4\mu_{\sigma^2_x}} >  0$$ {#eq-ineq_rearr1}

Secondly, as all terms contain $\mu_{\sigma^2_x}$ in the denominator, to
arrive at @eq-ineq_rearr2 we divided by $\mu_{\sigma^2_x}$.

$$\frac{\tau_{MD}^2}{R_1} - \tau_{MD}^2 + \frac{\mu_{MD}^2}{R_1^2 \mu^2_{\sigma^2_x}} \frac{R_2 \tau^2_{\sigma_x^2}}{4 R_1} - \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} \frac{\tau^2_{\sigma_x^2}}{4} >  0$$ {#eq-ineq_rearr2}

For @eq-ineq_rearr3 we properly combined the multiplied fractions.

$$\frac{\tau_{MD}^2}{R_1} - \tau_{MD}^2 + \frac{\mu_{MD}^2 R_2 \tau^2_{\sigma_x^2}}{4R_1^3 \mu^2_{\sigma^2_x}} - \frac{\mu_{MD}^2\tau^2_{\sigma_x^2}}{4\mu^2_{\sigma_x^2}}  >  0$$ {#eq-ineq_rearr3}

Taking $\frac{\tau_{\sigma^2_x}}{\mu_{\sigma^2_x}} = CV{\sigma^2_x}$
allows to simplify for @eq-ineq_rearr4.

$$\frac{\tau_{MD}^2}{R_1} - \tau_{MD}^2 + \frac{\mu_{MD}^2 R_2}{4R_1^3}CV^2{\sigma^2_x} - \frac{\mu_{MD}^2}{4}CV^2{\sigma^2_x}  >  0$$ {#eq-ineq_rearr4}

As a last step, factoring out finalizes rearranging the terms from
@eq-ineq_R to @eq-ineq_fin:

$$ \tau_{MD}^2\left(\frac{1}{R_1} - 1\right) + \frac{\mu_{MD}^2 CV^2_{\sigma^2_X}}{4} \left(\frac{R_2}{R_1^3}  - 1\right) > 0$$ {#eq-ineq_fin}

The circumstances under which the inequality described in @eq-ineq_fin
holds, are circumstances where the claims made in W&D and H&S are
directly contradicted, as in those cases the ES heterogeneity is larger
after applying an attenuation correction procedure. Concerning the terms
in @eq-ineq_fin, we know that all terms left-hand of the inequality,
outside of the brackets ($\tau^2_{MD}$, $\mu^2_{MD}$ and
$CV^2_{\sigma^2_X}$), are bound to be positive. For the inequality to
hold, we need the left-hand side of the equation to remain positive,
larger than zero. We can distill two scenarios, under which this
equation should hold: (a) one of the terms inside the brackets
($\frac{1}{R_1} - 1$ or $\frac{R_2}{R^3_1}  - 1$) is positive and large
enough, so that the second term not containing that same bracket is
positively dominated by the first term, meaning it is sufficiently large
that the positive term cancels out the negative term; or (b) both terms
inside the brackets need to be positive.

Generally, we know that both $R_1$ and $R_2$ are bound to be positive,
as both contain different, strictly positive parameters of the
distributions of true and observed score variance. Additionally, we know
that $R_1$ is equivalent to the average score reliability, and therefore
bound between $0\le R_1\le1$. Therefore, the term inside the first
bracket is bound to be positive ($\frac{1}{R_1} – 1$). Similarly, $R_2$,
as the ratio of true and error score variance heterogeneity, is bound
between $0\le R_2\le1$. If $R_1^3$ is larger than $R_2$, then the term
inside the second bracket turns negative. Generally, this means that the
inequality can only be violated if the relative remaining heterogeneity
of score variance ($R_2$) is smaller than the the average score
reliability to the power of 3. Restated in mathematical terms, this
leads us to the rule/condition described in @eq-ineq_short, in the main
text.

However, it is crucial to note that violation of $R_1^3 \leq R_2$ is
just a necessary, but not sufficient condition for the inequality not to
hold anymore. Should $R_1^3 \leq R_2$ not hold, the negative term in
@eq-ineq_fin still needs to be sufficiently large, that it can
negatively dominate the positive term. That would mean that the
combination of parameters $\mu_{MD}$, $\tau_{MD}$ and $CV_{\sigma^2_x}$
need to take on values, that allow for such a scenario. Since we do not
want to make any claims regarding the unstandardized mean difference and
how it relates to the coefficient of variation of score variance, we can
not distill a scenario in which it is guaranteed, that @eq-ineq_fin is
violated. Therefore $R_1^3 \leq R_2$ only describes a condition, under
which it is guaranteed that @eq-ineq_fin holds. If $R_1^3 \leq R_2$ is
violated, it is a first necessary condition for the entire inequality to
be violated, but it is not sufficient as it still depends on other
parameters.

### B - Examining bias due to delta-method

We want to examine whether applying the delta-method sufficiently
approximates the moments of the distribution of score variance to derive
the conditions under which the inequality @eq-ineq_init holds. To do so,
we compared estimates of mean and heterogeneity of total score variance
$\sigma^2_X$ derived using Boot-Err, and by running a meta-analysis on
total score standard deviation $\sigma_x$ and subsequently applying the
delta-method to estimate mean and heterogeneity of $\sigma^2_X$ .
Therein, we generated two sets of estimates of $\mu_{\sigma^2_x}$ and
$\tau^2_{\sigma^2_x}$, one via Boot-Err and one via the delta-method. If
there are systematic differences between the sets, this points towards
bias in the delta-method approximation. The estimates can be found in
@tbl-delta.

```{r, echo = FALSE}
#| label: tbl-delta
#| tbl-cap: Empirical assessment of appropriateness of delta-method
# knitr::kable(read.csv(here("Tables/delta_test.csv")),
#              col.names = c("MASC", 
#                            "$\\hat{\\mu}_{\\sigma^2_X}$", 
#                            "$\\hat{\\tau}^2_{\\sigma^2_X}$", 
#                            "$\\hat{\\mu}_{\\Delta \\sigma^2_X}$", 
#                            "$\\hat{\\tau}^2_{\\Delta \\sigma^2_X}$"))

gt::gt(read.csv(here("Tables/delta_test.csv")) %>% 
         arrange(unique.ES_rma_df.MASC..nn_eff_idx.)) %>% 
  gt::cols_label(unique.ES_rma_df.MASC..nn_eff_idx. = "MASC",
                 vars.mu_X = gt::md("$\\hat{\\mu}_{\\sigma^2_X}$"),
                 vars.tau2_X = gt::md("$\\hat{\\tau}^2_{\\sigma^2_X}$"),
                 delta_mu_X = gt::md("$\\hat{\\mu}_{\\Delta \\sigma^2_X}$"),
                 delta_tau2_X = gt::md("$\\hat{\\tau}^2_{\\Delta \\sigma^2_X}$")) %>% 
  gt::tab_options(table.width = gt::pct(100),
                  column_labels.font.weight = "bold") 
```

*Note:* $\hat{\mu}_{\sigma^2_X}$ is the back-transformed meta-analytic
estimate of mean total score variance, obtained via Boot-Err;
$\hat{\mu}_{\sigma^2_X}$ is the back-transformed estimate of its
heterogeneity. On the other hand, $\hat{\mu}_{\Delta\sigma^2_X}$ is the
back-transformed meta-analytic estimate of mean total score variance,
obtained from application of the delta method;
$\hat{\mu}_{\Delta\sigma^2_X}$ is the back-transformed estimate of its
heterogeneity.

Generally, for 11 out of 12 phenomena, we find miniscule, but mostly
negative, differences between the two estimation strategies. This means
that there is a slight downward bias in the estimates from the
delta-method. However, for the majority of phenomena, the differences
occur in the 2nd decimal, indicating that violations in the assumptions
concerning the delta-method are negligible here. Only for the phenomenon
described in PSACR002 (Behaviour) we find larger differences. These are
large enough to be concerning and pertain largely the estimate of
heterogeneity. Accordingly, inferences regarding this phenomenon need to
be interpreted with caution. Generally, since for the majority of
phenomena we find very little bias and as the delta-method is not used
to generate estimates beyond this particular section, we believe its use
to be defensible.