### 2025-12-23 Manuscript_MD_ScoreReLiability.qmd
Update - FR comments

---
title: "Differences in score reliability do not explain meta-analytic heterogeneity in standardised effect sizes"
bibliography: ReLiability_references.bib
csl: styles/apa-7th-edition.csl
format: docx
editor: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
# library loading and installing as necessary

# relevant libraries required for this script
packages <- c("ggplot2", "here", "metafor", "gridExtra", "knitr", "magrittr", "dplyr", "gt")

# check, whether library already installed or not - install and load as needed:
apply(as.matrix(packages), MARGIN = 1, FUN = function(x) {

  pkg_avail <- nzchar(system.file(package = x))   # check if library is installed on system

  if(pkg_avail){
    require(x, character.only = TRUE)             # load the library, if already installed

  }else{
    install.packages(x, repos = "https://ftp.gwdg.de/pub/misc/cran/")                           # install the library, if missing
    require(x, character.only = TRUE)             # load after installation
  }
})

source(here("ReLiability_Function-library.R"))

ES_rma_df <- read.csv(here("Data/Processed/Aggregates_ES_analysis.csv"))

agg_L <- readRDS(here("Data/Processed/Aggregates_simple.csv"))

MASC_names <- unique(ES_rma_df$MASC)

```

#

Lukas J. Beinhauer *University of Erfurt* ORCID:
https://orcid.org/0000-0001-9841-3089

Jens H. Fünderich *Ludwig Maximilian University of Munich* ORCID:
https://orcid.org/0000-0002-7185-9248

Frank Renkewitz *University of Erfurt* ORCID:
https://orcid.org/0000-0001-8072-6802

#### Abstract

Heterogeneity in standardized effect sizes, such as Cohen’s d, is
informative because it indicates that a psychological phenomenon is not
yet fully understood. Prior work by Hunter and Schmidt as well as
Wiernik and Dahlke has argued that such heterogeneity partly reflects
differences in measurement precision and that correcting for these
differences using attenuation correction should therefore reduce
heterogeneity. We reanalyze data from large-scale collaborative
projects, including the ManyLabs studies, Registered Replication
Reports, and the Psychological Science Accelerator, and demonstrate that
this assumption does not hold universally. Treating standardized effect
sizes as ratio variables, we offer an alternative account of how
differences in measurement precision relate to observed heterogeneity.
Across many psychological research contexts, score reliability tends to
be high and measurement error relatively stable. Under these conditions,
when samples differ substantially in their composition or diversity,
attenuation correction procedures are more likely to increase rather
than decrease heterogeneity. We conclude by discussing the limited
relevance of random measurement error for explaining effect size
heterogeneity in the context of convenience sampling.

#### Non-technical abstract

Researchers often find that studies on the same psychological effect
report different results, known as differences in effect sizes. Such
variation suggests that the phenomenon is not yet fully understood. Some
have argued that part of this variation comes from differences in
measurement precision—how reliably a study’s tools or tests measure what
they intend to measure. Based on this reasoning, it has been suggested
that adjusting or “correcting” for these differences should make results
more consistent across studies. We revisited this assumption by looking
at large collaborative projects in psychology, such as ManyLabs, the
Registered Replication Reports, and the Psychological Science
Accelerator. We found that the expected reduction in variation does not
generally occur. In fact, when measurement precision is already quite
high - as it usually is in psychology, where researchers carefully
select reliable measures - but samples differ widely in who
participates, such corrections can actually increase differences in
reported effect sizes. In simpler terms, differences in how diverse the
samples of participants are across studies (which comes about from
convenience sampling) matters more for explaining variation in results
than small differences in how precisely a measure captures a construct.
We introduce a simple rule that identifies when correcting for
measurement differences is likely to increase rather than reduce
heterogeneity. We conclude by discussing why random measurement error
plays a smaller role in explaining effect size differences than the use
of diverse, convenience-based samples.

Large-scale collaborative replication efforts have sparked discussions
surrounding the replicability of psychological phenomena. For attempts
to estimate a phenomenon’s replicability or to predict whether a future
replication will replicate successfully, heterogeneity is a crucial
parameter that needs to be assessed [@stanley2018]. In the meta-analytic
context, heterogeneity describes the variation of effect sizes, free of
sampling error [@dersimonian1986; @higgins2002; @riley2011]. Therefore,
the presence of heterogeneity can imply that some replications of a
single phenomenon may be successful, while others are not. If
heterogeneity increases, meaning the phenomenon’s effect size varies
more strongly for unexplained reasons, the probability of observing an
effect size around zero or even in the negative space grows larger as
well [@kenny2019]. If we know the mean size and heterogeneity of a
phenomenon’s effect, we would theoretically be able to establish an
expected replication rate [e.g. @vu2024].

Similarly, it has been argued that heterogeneity in effect sizes is an
indicator of the theory’s “completeness” surrounding the phenomenon. For
example, Linden and Hönekopp argue that “low (as opposed to high)
heterogeneity reflects a more advanced understanding of the subject
matter being studied” [-@LindenHönekopp, p. 2]. Similarly, von Hippel
and Schuetze [-@vonHippel2023] argue that heterogeneity in effects is an
indicator of a vague, poorly specified theory.

Large-scale attempts of direct replications, using identical protocols,
such as the Many Labs studies or Registered Replication Reports
[@ManyLabs1; e.g. @RRRHart], for the first time allow researchers to
estimate heterogeneity undistorted by selection processes like
publication bias. In re-analyses of these studies, Olsson-Collentine et
al. [-@olsson2020] identify a positive correlation between a
phenomenon’s effect size and its heterogeneity. Both van Erp et al.
[-@vanErp2017] and Stanley et al. [-@stanley2018] estimate strong
degrees of heterogeneity across conceptual replications in Psychology.
In a separate re-analysis of large scale direct replications, Renkewitz
et al. (in preparation) identify substantial heterogeneity in almost all
projects where a non-zero effect could be identified. This aligns with
the correlation found by Olsson-Collentine et al. [-@olsson2020].

The search for moderators that might explain said heterogeneity across
several phenomena has not been particularly successful [e.g. @ManyLabs3,
@ManyLabs2]. Another source for this heterogeneity, which is not a
typical moderator tied to a specific theory, might be impairments in
score reliability which vary in their extent across replications
[@olsson2020]. Hunter and Schmidt [-@hunter2004] and Wiernik and Dahlke
[-@wiernik2020], discussing the role of score reliability for
standardized effect sizes, claim that differences in score reliability
should inflate, and therefore increase, heterogeneity. This implies that
standard corrections for imperfect score reliability, such as
attenuation correction procedures, should remove that inflation,
reducing heterogeneity in corrected effect sizes.

Over the following pages, we assess whether score reliability can in
fact be made responsible for heterogeneity identified across several
psychological phenomena. However, we find no empirical evidence for the
argument that correcting for differences in score reliability reduces
effect size heterogeneity. Instead, we find that heterogeneity might as
well be larger after the correction procedure. Therefore, previous
analyses and assumptions regarding the relationship between score
reliability and heterogeneity need to be false. In the second half of
this text we make use of analytical arguments to identify the necessary
conditions under which an attenuation correction inevitably leads to an
increase in effect size heterogeneity.

## Effect sizes and score reliability in meta-analysis

In both the initial reports of large scale replication attempts, as well
as the re-analyses by Olsson-Collentine et al. [-@olsson2020],
standardized effect sizes (from here-on abbreviated as ES) such as
Cohen’s d or Hedge’s g were used. For the remainder of the article,
Cohen’s d, defined in @eq-d, will be used as an exemplary estimate of
ES, as it is used across a wide range of contexts and well understood by
a broad audience.

$$d = \frac{MD}{\sigma_X}$$ {#eq-d}

Here, MD refers to the mean difference between two groups of interest,
while $\sigma_X$ is the pooled standard deviation. As is widely known,
score reliability affects ES [e.g. @hunter2004; @wiernik2020]. In the
context of classical test theory (CTT) score reliability $\rho_{XX’}$ is
defined as the ratio of true to total score variance [@lord2008], as
defined in @eq-rel.

$$\rho_{XX’} = \frac{\sigma^2_T}{\sigma^2_X}=\frac{\sigma^2_T}{\sigma^2_T + \sigma^2_E}$$ {#eq-rel}

In this equation $\sigma_T^2$ refers to the true score variance in the
sense of CTT, meaning the actual variance of the variable, undistorted
by random measurement error. In the same sense, $\sigma_X^2$ refers to
the total variance of scores, including both the true variance and the
random error variance $\sigma_E^2$. This implies that, if true score
variance is assumed to be constant across replications, a lower score
reliability can only occur due to a larger error score variance. Total
score variance would also be larger, therein leading to a smaller ES, as
opposed to a similar data-set where a higher score reliability is
achieved through a smaller error score variance. However, score
reliability is an aspect of a measuring instrument applied to a
population [@lord2008; @vacha1998]. In the meta-analytic context, as
studies tend to be replicated in a variety of populations, neither true
nor error score variance can be expected to always be identical across
replications. This most likely leads to heterogeneity in score
reliability, which, as discussed, is bound to affect ES heterogeneity as
well.

Previous discussions of heterogeneity in score reliability have
exclusively discussed it as a parameter that inflates heterogeneity in
ES, implying that, if score reliability was perfect across all
replications, the actual heterogeneity would have been lower. In their
discussion, Wiernik and Dahlke claim that „*Measurement error variance
will impact the results of meta-analyses in three ways: by (a) biasing
the mean effect size toward zero, (b) inflating effect-size
heterogeneity and confounding moderator effects, and (c) confounding
publication-bias and sensitivity analyses*” [p. 3, -@wiernik2020].
Additionally, more clearly, they state that “*If the studies included in
a meta-analysis differ in their measures’ reliabilities, heterogeneity
estimates will be artifactually inflated, erroneously suggesting larger
potential moderator effects*” [p. 4, -@wiernik2020]. They base their
claims largely on work done by Hunter and Schmidt [-@hunter2004], who
claim that “*Variation in reliability across studies causes variation in
the observed effect sizes above and beyond that produced by sampling
error.*” (p. 302). Overall, both references imply that differences in
score reliability inevitably lead to an inflated heterogeneity in ES.

As both sources Hunter and Schmidt [-@hunter2004] and Wiernik and Dahlke
[-@wiernik2020] will be referred to repeatedly, we abbreviate these
references as H&S and W&D respectively.

### Attenuation correction

If information concerning score reliability is available, it is possible
to correct the individual ES for its unreliability. This process is also
known as attenuation correction and, while not without its criticisms
[@muchinsky1996; @schmidt1996; @schennach2016], is a repeatedly
recommended practice in Psychology [@hunter2004; @wiernik2020]. @eq-dc
describes a simple attenuation correction procedure.

$$d_c = \frac{d}{\sqrt{\hat{\rho}_{xx’}}}$$ {#eq-dc}

Here, $d_c$ describes the attenuation-corrected estimate of $d$, which
is corrected by dividing it by the estimate of score reliability
$\hat{\rho}_{xx’}$. As described in W&D or Lord & Novick [-@lord2008],
this is essentially the same, as if the corrected estimate of $d_c$ was
constructed using an estimate of true score standard deviation, as in
@eq-dct.

$$d_c = \frac{MD}{\hat{\sigma}_{T}}=\frac{MD}{\sqrt{\hat{\rho}_{xx’} \hat{\sigma}_X^2}}$$ {#eq-dct}

$\hat{\sigma}_T$ can be estimated by simply rearranging the terms from
@eq-rel. Thereby, an estimate of ES is generated, under the premise that
score reliability would have been perfect ($\rho_{xx'} = 1$).

The claims made in W&D imply that corrections of individual observed ESs
for their unreliability should lead to lower estimates of heterogeneity
in meta-analyses of corrected ES. If heterogeneity in score reliability
adds to ES heterogeneity, attenuation correction procedures should
eliminate that additional variation, as all corrected ES come with
identical (perfect) score reliability.

Re-analyses of the collaborative multi-site replication efforts have
uncovered differences in score reliability across replications of the
same phenomena [@shaw2020]. The claims made in W&D and H&S imply that
the ES heterogeneity uncovered in these projects therefore could
potentially be explained by these differences in score reliability.
Similarly, in a re-analysis of multi-site replication projects,
Olsson-Collention et al. [-@olsson2020] discuss differences in score
reliability as a potential moderator for ES heterogeneity.

Therefore, we assess the extent to which this intuition is supported by
empirical evidence in the following section. The re-analysis of archival
data serves to shed light on the following research question:

-   To what extent can heterogeneity in standardized effect sizes be
    reduced by means of an attenuation correction?

## Re-analysis of archival data

We have collected a large number of openly available multi-site direct
replications on psychological phenomena [@dripht2024]. The data-sets of
about 55 phenomena, largely stemming from the efforts of the Many Labs
studies, Registered Replication Reports or the Psychological Science
Accelerator, are made openly available in a standardised format at
\[<https://osf.io/7p8rk/>\], the DRIPHT repository. A documentation of
the standardised format, including a web-application to easily explore
the public data-sets, can be found in Fünderich et al.
[-@fuenderich_metapipex2024].

### Data

The ManyLabs projects were collaborative efforts to replicate several
psychological phenomena across different research sites, employing
identical protocols. From five published projects, the data for the
first three and the fifth Many Labs projects [@ManyLabs1; @ManyLabs2;
@ManyLabs3; @ManyLabs5] were harmonized and publicly stored in a
standardized format when the DRIPHT repository was set up [@dripht2024].
The Registered Replication Reports are similar collaborative efforts,
with the sole distinction that for each report a single phenomenon was
replicated several times. The Registered Reports 3-10 were added to the
DRIPHT repository, as they employed experimental designs with at least
two groups [@RRRHart; @RRRSripada; @RRRFinkel; @RRRStrack; @RRRRand;
@RRRDijksterhuis; @RRRSrull; @RRRMazar]. Lastly, some projects from the
Psychological Science Accelerator were added to the DRIPHT repository,
namely PSA006, PSACR001, PSACR002 and PSACR003 [@PSA006; @PSACR001;
@PSACR002; @PSACR003]. In contrast to Many Labs or Registered
Replication Projects, these collaborative efforts, while also
distributed across the globe, do not formally attempt to perform direct
replications. Instead, these projects focus on original research or
conceptual research done collaboratively in laboratories across the
globe. However, as each project used the same protocol and participants
are distributed across different countries, a data structure similar to
that of a multi-site replication study emerges. What makes all these
collaborative efforts valuable for this manuscript is the fact that all
phenomena collected in the DRIPHT-repository are making use of two-group
designs and are therefore easily assessed using standardised mean
differences.

While replication data on a decent number of psychological phenomena is
available in the DRIPHT-repository, selected phenomena need to fulfil
three conditions to demonstrate how incorporating score reliability
affects heterogeneity in ES: (1) As we focus on the use of ES $d$, the
phenomenon needs to be assessed using a group design consisting of at
least two groups. (2) Score reliability estimates can be derived, this
implies that the phenomenon needs to be measured using either several
indicators forming a single scale or by repeatedly measuring across
several timepoints. (3) Score reliability can only attenuate effect
sizes that are not zero in the first place. Therefore, we focus on
phenomena where meta-analytic mean ES can be statistically distinguished
from zero. This aligns with the observation made in (Renkewitz,
Fünderich & Beinhauer, in preparation) and Olsson-Collentine et al.
[-@olsson2020]. As null-effects displayed little to no heterogeneity,
there is no variability that might be reduced by incorporating
differences in score reliability. While the first criterion is fulfilled
by all phenomena found in the DRIPHT repository, the second and third
criteria still need evaluation.

As all phenomena in the collection fulfil the first condition, 50
phenomena have been catalogued where the effect is studied by comparing
the outcome across two separate groups. From the collection of 50
phenomena, 21 fulfil the second condition, so that estimates of score
reliability can be derived. For the 21 phenomena, the dependent variable
used to construct the effect size is measured by employing more than a
single indicator. The number of indicators per scale varied between 3
and 30. The majority of scales was made up of 3 to 6 items. At the same
time, however, this implies that the remaining 29 phenomena did not make
use of more than a single indicator to measure the dependent variable.
Therefore, for the majority of phenomena it is not possible to assess
whether measurement quality is sufficiently high in terms of internal
consistency or low random error variance.

From the 21 phenomena that employed more than a single indicator, 19
made use of Likert-style items, where respondents would indicate their
agreement to some kind of statement. The scale-length varied from 3 to
10-points. The items measuring the remaining 2 phenomena were coded
dichotomously, as responses were right or wrong. While the majority of
designs make use of some form of control-treatment manipulation,
randomly assigning participants to a condition (19 phenomena), some
phenomena refer to the effect of pre-existing differences, e.g.
biological sex (2 phenomena). Subsequently, those phenomena need to be
identified where the meta-analytic mean effect size is statistically
significant different from zero. This will be done as a preliminary step
in this manuscript’s analysis procedure. Detailed information concerning
the different phenomena and how they were measured can be found at
\[https://osf.io/g2fzq\], while a brief summary can be found in appendix
C.

To assess whether ES heterogeneity can be reduced by correcting for
score reliability, estimates of meta-analytic heterogeneity are compared
across two situations: first, heterogeneity of raw ES, as computed in
@eq-d is assessed. Secondly, ES are corrected for imperfect reliability.
Computing the heterogeneity of these corrected ES allows for an
assessment, whether the heterogeneity did indeed shrink compared to the
first situation, as predicted in W&D and H&S.

In order to derive estimates of meta-analytic heterogeneity, we make use
of a random-effects meta-analysis. Broken down, a random-effects
meta-analysis essentially pools the sampling uncertainty of the
different estimates that are to be aggregated. By removing this pooled
estimate from the observed variability in the effect size estimates, we
can generate an estimate of its heterogeneity - how strongly the effect
sizes vary beyond their pooled sampling uncertainty.

For the standardized mean difference $d$, we can compute the standard
error as an estimate of its sampling uncertainty. The standard error of
Cohen’s $d$ is computed as defined in @eq-SEd.

$$SE_d=\sqrt{\frac{n_1+n_2}{n_1n_2}+\frac{d^2}{2\left(n_1+n_2\right)}}$$ {#eq-SEd}

As the corrected estimate of Cohen’s $d$ is larger and requires an
additional unknown parameter, an estimate of score reliability, the
larger uncertainty in this parameter should be reflected in a larger
standard error. The standard error for $d_c$ is defined in @eq-SEdc.

$$SE_{d_c} = \sqrt{SE^2_d \frac{d_c}{d}}$$ {#eq-SEdc}

The estimates concerning the meta-analytic mean and heterogeneity of
these standardized effect sizes attempt to describe their population
distribution. Therefore, over the following paragraphs, whenever we
refer to estimates derived by means of random effects meta-analysis, we
will use the Greek letter $\delta$ to denote estimates of uncorrected
$d$'s population distribution and $\delta_c$ to denote estimates of
corrected $d_c$'s population distribution.

### Methods

For each replication of a phenomenon, Cohen’s $d$ is computed as an
estimate of raw ES, according to @eq-d, including its standard error
(@eq-SEd). Subsequently, a random-effects meta-analysis is performed
using metafor version 4.8-0 [@metafor]. To estimate meta-analytic mean
ES and heterogeneity we make use of the REML-estimator, as that
estimator tends to perform best across a range of conditions
[@hoenekopp2022]. To identify which phenomena pass criterion (3) we use
a Wald-type significance test, assessing for which phenomenon the
meta-analytic mean ES is significantly different from zero. For this
hypothesis test, as all other hypothesis tests in this manuscript, we
employ a one-tailed significance level of .05.

Additionally, estimates of score reliability are derived. While not
without criticism [e.g. @sijtsma2009], Cronbach’s Alpha is used to
estimate score reliability. It is often noted that tau-equivalence is an
unrealistic assumption to hold against real-world data, implying that
other estimates of score reliability, such as McDonald’s Omega
[@mcdonald1999], Guttman’s Lambda 2 or 4 [@guttman1945], or the Greatest
Lower Bound [@tenBerge2004] might be better suited. However, all scales
employed in this project are scored by computing the simple mean or sum
of responses across items for each individual. Computing the mean or sum
of several items implicitly assumes tau-equivalence, as each item
contributes equally to the individual’s test score. Therefore,
Cronbach’s Alpha is the better choice to estimate score reliability, as
it avoids a mismatch in assumptions between how the score is computed
and how score reliability is estimated. The individual estimates of
Cohen’s $d$ are corrected using estimates of Cronbach's Alpha, according
to @eq-dc, including the corrected standard errors (@eq-SEdc). Thereby,
estimates of ES are computed, which are corrected for imperfect
reliability. Again, a random-effects meta-analysis is run using metafor,
generating an estimate of heterogeneity in corrected ES. Different
indicators of heterogeneity are readily available, such as $I^2$, $H^2$
[@borenstein2009; @higgins2002]. A relative heterogeneity measure less
common in Psychology, but easily computed is the coefficient of
variation $CV = \frac{\tau}{\mu}$. However, the descriptions found in
H&S and W&D discuss the absolute amount of heterogeneity in terms of
variance ($\tau^2$) or standard deviation ($\tau$). Therefore, we also
make use of absolute heterogeneity here.

Subsequently, in order to assess whether ES heterogeneity was indeed
reduced by the attenuation correction procedure, the estimates of
heterogeneity in uncorrected ES and corrected ES are compared. Using
Cochran’s *Q*-test [@cochran1954], we assess whether the estimates of
heterogeneity are statistically significantly different from 0. Only for
those projects where we have sufficient confidence that the individual
estimates of heterogeneity are larger than zero can we actually begin to
interpret whether there was any change.

To identify whether a reduction in ES heterogeneity is indeed
accompanied by differences in score reliability, the estimates of score
reliability are assessed meta-analytically as well. To do so, a
Reliability Generalization Meta-Analysis is performed [@vacha1998;
@sanchez2013]. This entails that individual estimates of score
reliability are adequately transformed using Bonett’s transformation:

$$T_{\hat{\rho}} = ln(1 - \hat{\rho})$$ {#eq-Bonett}

The transformation in @eq-Bonett has variance-stabilising properties and
therefore allows for adequate inferences concerning heterogeneity in
score reliability [@bonett2002]. Cochran’s Q-test is used to identify
statistically significant heterogeneity in transformed score
reliability. Estimates derived from a Reliability Generalization
Meta-Analysis using these transformations can be back-transformed to the
original score reliability scale [@beinhauer2025]:

$$\mu_{\rho{xx'}} \approx 1 - exp(\mu_{T\left[\rho_{XX^\prime}\right]})-\frac{1}{2} exp(\mu_{T\left[\rho_{xx'}\right]})\tau_{T\left[\rho_{XX^\prime}\right]}^2$$ {#eq-rel_back_mu}

In @eq-rel_back_mu, $\mu_{\rho_{XX'}}$ is the back-transformed
meta-analytic mean score reliability - on the original coefficient's
scale between 0 and 1. Similarly, $\tau_{\mu_{\rho_{XX'}}}$ is the
back-transformed heterogeneity in score reliability:

$$\tau_{\rho_{xx'}}^2 \approx exp(\mu_{T\left[\rho_{xx'}\right]})2 \tau_{T\left[\rho_{xx'}\right]}^2+exp(\mu_{T\left[\rho_{xx'}\right]})^2\tau_{T\left[\rho_{xx'}\right]}^4$$ {#eq-rel_back_tau}

We executed these steps for each phenomenon. The statistical programming
language R, Version 4.4.2 [@RCore] is used for all data manipulation and
statistical analysis. Graphics are constructed using the R package
*ggplot2*, version 3.5.1 [@ggplot2].

### Results

[Equations @eq-d], [-@eq-dc], [-@eq-SEd], and [-@eq-SEdc] were used to
generate estimates of standardized ES, both uncorrected and corrected,
with their corresponding standard errors. Generally, this leads to
larger (absolute) ES and larger standard errors after the correction. As
an example from the 21 selected phenomena, the results from this
procedure on replications of Nosek’s study on sex differences in
preferences for the subject of Arts is displayed in the forest plot in
@fig-forest. Here, grey dots represent the uncorrected ES for each
sample, while black dots represent the corrected ES in each sample. The
bars surrounding these dots show the respective 95%-CIs.

```{r, echo = FALSE, fig.height = 6}
#| label: fig-forest 
#| fig-cap: Forest Plot ML1 - replications of Nosek's study on sex differences in subject preferences (Arts)
forest_plot_rel(ES_rma_df[which(ES_rma_df$MASC == "Nosek_Explicit_Art"),], agg_L[[which(MASC_names == "Nosek_Explicit_Art")]])
```

*Note:* Grey dots represent uncorrected estimates of $d$, while black
dots represent corrected estimates of $d_c$. Similarly, grey bars
visualise the 95%-CI for uncorrected estimates, black bars the 95%-CIs
for corrected estimates. The grey diamond represents the meta-analytic
estimate of uncorrected ES $\mu_{d}$, its width represents the 95%-CI of
said estimate. The black diamond represents the same parameters for the
meta-analytic estimate of corrected ES $\mu_{d_c}$,.

In @fig-forest it becomes apparent that the reliability attenuation
procedure leads to an increase in the individual absolute effect size,
as the black dots are moved further from zero. Simultaneously, the
standard errors grew larger after the attenuation correction, which
leads to larger 95%-CIs. This is most easily observed for the
MTurk-sample, which had a rather large confidence interval to begin
with. Additionally, in the diamond at the bottom of the figure, we can
see that the meta-analytic mean ES is also larger after the attenuation
correction took place. @tbl-meanES contains estimates and test results
for all phenomena, both corrected and uncorrected.

#### Which phenomena pass criterion (3)?

@tbl-meanES demonstrates that what we observed in @fig-forest holds
across all 21 phenomena. All meta-analytic effects are larger in their
absolute size after applying an attenuation correction procedure.
Similarly, the uncertainty quantified in the estimate's standard error
is larger. Additionally, @tbl-meanES highlights which phenomena pass
criterion (3)—the effect size must be statistically significantly
distinguishable from zero at $p < .05$. Twelve phenomena pass this
criterion. It may be remarked that the attenuation correction procedure,
in this data, appears to have no influence on whether a phenomenon
passes the criterion. The differences in p-value are rather small and
lead to no difference in conclusion, regarding a significance level of
5%.

```{r, echo = FALSE}
#| label: tbl-meanES
#| tbl-cap: Meta-analytic average  ES

gt::gt(read.csv(here("Tables/Mean_ES.csv")) %>% 
         arrange(MASC)) %>% 
  gt::cols_label(MASC = "MASC",
                 mu_str_raw = gt::md("$\\hat{\\mu}_{d}$ (SE)"),
                 #mu_str_raw = gt::md("Mean $d$ (SE)"),
                 pval_raw = "p",
                 mu_str_cor = gt::md("$\\hat{\\mu}_{d_c}$ (SE)"),
                 #mu_str_cor = gt::md("Mean $d_c$ (SE)"),
                 pval_cor = "p") %>% 
  gt::tab_options(table.width = gt::pct(100),
                  column_labels.font.weight = "bold") # %>% 
  # gt::cols_width(pval_raw ~ gt::px(40),
  #                pval_cor ~ gt::px(40))
```

*Note:* In this table $\hat{\mu}_{d}$ represents the meta-analytic
estimate of uncorrected $d$ and $\hat{\mu}_{d_c}$ represents the the
meta-analytic estimate of corrected $d_c$.

#### Estimates of heterogeneity

@tbl-tauES contains the results from performing an RG-MA on the
Cronbach's Alpha coefficients of the scales used to measure the 12
remaining phenomena. Additionally, it contains estimates of
heterogeneity for both uncorrected ES ($d$) and corrected ES ($d_c$),
which are central to assess the research question posed in section
*Attenuation Correction*. Most importantly, @tbl-tauES demonstrates that
for all but one of the phenomena, the attenuation correction procedure
barely has an impact on the extent of ES heterogeneity. However, while
the heterogeneity for the phenomenon from RRR9 (Behaviour) seems to be
reduced by about a third of its initial value, the ES heterogeneity for
this phenomenon is rather small to begin with. Using a significance
test, we can not distinguish this value confidently from zero. This
implies that for none of phenomena, differences in score reliability can
be made responsible for heterogeneity in standardized effect sizes.

```{r, echo = FALSE}
#| label: tbl-tauES
#| tbl-cap: Tests for ES Heterogeneity
#knitr::kable(x = read.csv(here("Tables/Heterogeneity_ES.csv")))

tab_hetES <- read.csv(here("Tables/Heterogeneity_ES.csv"))
tab_hetrel <- read.csv(here("Tables/Results_RMA_alpha.csv"))

tab_both <- data.frame(tab_hetES$MASC,
                       tab_hetrel[,c("mu_alpha_str", "tau_alpha", "QE_str", "QEp")],
                       tab_hetES[,c("tau_raw", "QE_raw_str", "QEp_raw", "tau_cor", "QE_cor_str", "QEp_cor")])

# knitr::kable(tab_both,
#              col.names = c("MASC", 
#                            "$\\hat{\\mu}_{\\rho_{XX'}}$ (95%-CI)", 
#                            "$\\hat{\\tau}_{\\rho_{XX'}}$", 
#                            "QE (df)", "p", 
#                            "$\\hat{\\tau}_{\\delta_0}$", 
#                            "QE (df)", "p",
#                            "$\\hat{\\tau}_{\\delta}$", 
#                            "QE (df)", "p"),
#              escape = FALSE)

gt::gt(tab_both  %>% 
  arrange(tab_hetES.MASC)) %>% 
  gt::cols_label(tab_hetES.MASC = "MASC",
                 mu_alpha_str = gt::md("$\\hat{\\mu}_{\\rho_{XX'}}$"),
                 tau_alpha = gt::md("$\\hat{\\tau}_{\\rho_{XX'}}$"),
                 QE_str = "QE (df)",
                 QEp = "p",
                 tau_raw = gt::md("$\\hat{\\tau}_{d}$"),
                 QE_raw_str = "QE (df)",
                 QEp_raw = "p",
                 tau_cor = gt::md("$\\hat{\\tau}_{d_c}$"),
                 QE_cor_str = "QE (df)",
                 QEp_cor = "p") %>% 
  gt::tab_options(table.width = gt::pct(100),
                  column_labels.font.weight = "bold") #%>% 
  # gt::cols_width(pval_raw ~ gt::px(40),
  #                pval_cor ~ gt::px(40))

# tab_one <- data.frame(tab_muES,
#                       tab_hetrel[,c("mu_alpha_str", "tau_alpha", "QE_str", "QEp")])
```

*Note:* $\hat{\mu}_{\rho_{XX'}}$ is the back-transformed meta-analytic
estimate of mean score reliability; $\hat{\tau}_{\mu_{\rho_{XX'}}}$ is
the back-transformed estimate of heterogeneity in score reliability;
$\hat{\tau}_{d}$ represents the estimate of heterogeneity in uncorrected
$d$; $\hat{\tau}_{d_c}$ represents the estimate of the heterogeneity in
corrected $d_c$.

Secondly, in @tbl-tauES we find that for ten out of twelve phenomena, ES
heterogeneity grows as a result from the attenuation correction
procedure. However, it needs to be noted that these differences are
rather small. For this reason, we do not mean to interpret these
differences, while we note that the direction of differences are clearly
inconsistent with expectations formulated in the literature discussed so
far. These results are incompatible with claims made in W&D and H&S.

Generally, however, it is noteworthy that removing differences in score
reliability only had a very little impact on ES heterogeneity. It
appears that, across the phenomena assessed in this project, the
heterogeneity before and after the correction procedure barely changes.
At least for the projects re-analysed here, differences in score
reliability do not affect ES heterogeneity as strongly as assumed.

Concerning the remaining two phenomena, for the phenomenon described in
ML5 - Shnabel (RPP), the reduction in heterogeneity is miniscule and the
estimate of heterogeneity is not statistically significant in the first
place. Most likely, the difference in ES heterogeneity estimates can be
explained by issues of estimation, as only eight replications are
available. Similarly, only for the phenomenon described in RRR9
(Behaviour), we find a seemingly substantial reduction in heterogeneity
due to the attenuation correction procedure. The estimates of ES
heterogeneity are not statistically significant, however, which hampers
our ability to meaningfully interpret and compare these estimates. For
the majority of phenomena it appears that differences in score
reliability do not inflate heterogeneity in uncorrected effect sizes.
Instead these differences mask ES heterogeneity that can be unveiled by
accounting for score reliability.

## Relationship between score reliability and effect size heterogeneity

In the re-analysis of archival data in the previous section we
demonstrated that, at least for the multi-site direct replication and
PSA-projects, score reliability does not hold any noteworthy value as a
potential moderator explaining ES heterogeneity. Additionally, we
demonstrated that claims formulated in W&D and H&S regarding how score
reliability heterogeneity increases ES heterogeneity are false. We found
a large number of phenomena where, contrary to popular claims,
correcting for score reliability increases ES heterogeneity, instead of
decreasing it.

This implies that there is an error in the reasoning or intuition behind
claims made in H&S and W&D, as our observations can not be explained by
the current literature. Over the following paragraphs, we will briefly
explore the intuition that might have led to the previous claims.
Subsequently, we will use analytical arguments to clarify how score
reliability actually affects ES heterogeneity. Therein, we also identify
potential conditions under which an attenuation correction will actually
increase ES heterogeneity - and under which it is more likely to
decrease it.

### Intuitive relationship

It might help to think about linear models, when discussing the
intuitive idea that controlling for score reliability should reduce ES
heterogeneity. Heterogeneity describes unexplained variation in a
parameter's true distribution, that we model using meta-analytic
techniques [@borenstein2009; @higgins2002]. In that way, the idea of
heterogeneity is closely related to residual variance, as we observe it
in a simple linear model. Both describe variation that could not be
explained by the model specifications, be it a model that employs
predictive variables like a simple linear model or a meta-analytic model
incorporating uncertainty in terms of standard errors.

The residual variance that remains in a simple linear model is reduced
in all instances, where we use a predictor that we know shares variance
with the predicted parameter. Generally, removing a source of variance
by controlling for it, we expect any unexplained variation to shrink.
Similarly, we know that differences in score reliability, by definition,
are related to differences in standardized effect sizes. Therefore,
intuitively, it seems sensible to expect that controlling for
differences in score reliability should reduce the unexplained variance
in standardized effect sizes, as described in H&S and W&D.

Hunter & Schmidt have additionally derived an equation which
demonstrates how heterogeneity in score reliability affects
heterogeneity in ESs [-@hunter2004, p. 309]. @eq-HS implies that
heterogeneity in $\delta$ is essentially a function in which
heterogeneity and mean value of $\delta_c$ are weighted by mean score
reliability $\rho_{xx‘}$ and its heterogeneity. Here, we adjusted the
notation from Hunter & Schmidt [-@hunter2004] to the notation used
through this text.

$$\tau^2_{\delta} = [\mu_{\rho_{xx‘}}]^2 \tau^2_{\delta_c} + [\mu_{\delta_c}]^2 \tau^2_{\rho_{xx‘}}$$ {#eq-HS}

Here, $\delta$ refers to the uncorrected ES, undistorted by sampling
error but not corrected for measuring error. With the index $c$,
$\delta_c$, on the other hand, refers to the corrected ES, undistorted
by sampling error and measurement error. Lastly, $\rho_{XX'}$ refers to
score reliability. Hunter & Schmidt [-@hunter2004] claim that @eq-HS
demonstrates that differences in score reliability inflate ES
heterogeneity, which means that an attenuation correction procedure
would have to reduce ES heterogeneity. However, it can be easily
verified that @eq-HS does not allow for such claims. Depending on both
the mean value $\mu_{\rho_{xx‘}}$ and the heterogeneity
$\tau^2_{\rho_{xx‘}}$ in score reliability, ES heterogeneity might be
increased or decreased by the differences in score reliability.

Most likely, Hunter and Schmidt did not fully explore the implications
of @eq-HS, but followed the general intuition we lined out – "If a
source of variation is controlled for, the amount of unexplained
variance should be reduced." Wiernik and Dahlke [-@wiernik2020] refer to
the work by Hunter & Schmidt [-@hunter2004], likely following their
intuition.

In the re-analysis of archival data, we have demonstrated that this
intuition does not hold for the relationship of ES heterogeneity and
score reliability. Instead, we propose an alternative way that helps us
discern under which conditions we can expect an attenuation correction
procedure to lead to an increase in ES heterogeneity.

### Describing ES as a random ratio variable

Realising that an ES, as defined in @eq-d, is actually a ratio, allows
for an analytical description alternative to @eq-HS of how differences
in score reliability can affect ES heterogeneity. If the phenomenon is
heterogeneous and score reliability varies across replications, it seems
sensible to assume that both numerator and denominator (MD and
$\sigma_X$) from @eq-d vary across replications. In that case, the ES
can be described as a random variable stemming from a ratio
distribution. A ratio distribution is a probability distribution
constructed by dividing one random variable by a second random variable:
$Z = \frac{X}{Y}$. First order Taylor approximation may be used to
generate estimates of mean and variance of the ratio variable
[@vanKempen1999]. Taylor approximations typically require no specific
assumptions regarding the distribution form, but that the variable (here
a ratio) is a) differentiable, b) has existing moments and that c) its
higher-order moments are negligible [e.g. @vandervaart2000]. Here we use
the Taylor-estimator of the ratio’s variance to demonstrate how
differences in score reliability may affect heterogeneity in ES.

$$\tau^2[Z] \approx \frac{\tau^2[X]}{\mu[Y]^2} - \frac{2\mu[X]}{\mu[Y]^3} cov[X,Y] + \frac{\mu[X]^2}{\mu[Y]^4} \tau^2[Y]$$ {#eq-XYZ_cov}

Assuming that the random variables are uncorrelated, simplifies the
equation to

$$\tau^2[Z] \approx \frac{\tau^2[X]}{\mu[Y]^2} + \frac{\mu[X]^2}{\mu[Y]^4} \tau^2[Y]$$ {#eq-XYZ}

Substituting for ES, unstandardized mean difference MD and pooled
standard deviation $\sigma_X$ leads to

$$\tau^2[\delta] \approx \frac{\tau^2[MD]}{\mu[\sigma_x]^2} + \frac{\mu[MD]^2}{\mu[\sigma_x]^4} \tau^2[\sigma_x]$$ {#eq-ratio_d0}

As stated in @eq-dct, correcting ES for score reliability can be
described through a correction of the pooled standard deviation:
$\sigma_T = \sqrt{\rho_{xx’} \sigma^2_x}$. Since $\rho_{xx'}$ is always
between zero and one, in the case of imperfect score reliability,
$\sigma_T$ will always be smaller than total $\sigma_x$.

In @eq-rel we see that total standard deviation $\sigma_X$ is
constructed only from error score variance $\sigma_E^2$ and true score
variance $\sigma_T^2$. Measurement precision should generally only
affect the error score variance. Therefore, if the heterogeneity in
$\sigma_X$ was introduced purely by differences in measurement
precision, an attenuation correction procedure can remove that
heterogeneity. Alternatively, the underlying latent variable may not be
distributed identically across replications. In that case, true score
variance $\sigma^2_T$ would vary across replications and induce
heterogeneity in $\sigma_X$. This heterogeneity would lead to
differences in standardized differences ES as well, but it can not be
removed by means of an attenuation correction procedure.

In that case, some heterogeneity in $\sigma_T$ would remain, albeit
heterogeneity found in $\sigma_X$ was reduced. @eq-ratio_d0 essentially
demonstrates how mean and heterogeneity of MD, paired with mean and
heterogeneity of total score standard deviation $\sigma_X$ can be used
to estimate heterogeneity in raw ES $\delta$. Similarly, we can adjust
@eq-ratio_d0 to describe how, instead of total score standard deviation,
mean and variance in true score standard deviation $\sigma_T$ affect
heterogeneity in corrected ES $\delta_c$.

$$\tau^2[\delta_c] \approx \frac{\tau^2[MD]}{\mu[\sigma_T]^2} + \frac{\mu[MD]^2}{\mu[\sigma_T]^4} \tau^2[\sigma_T]$$ {#eq-ratio_d}

The differences between the two [Equations @eq-ratio_d0] and
[-@eq-ratio_d] demonstrate how ES heterogeneity is affected by an
attenuation correction procedure. According to the quotes in H&S and
W&D, @eq-ratio_d0 should lead to a larger value of
$\tau^2\left[\delta\right]$, compared to the estimate of
$\tau^2\left[\delta_c\right]$ from @eq-ratio_d. However, as demonstrated
in the re-analysis of archival data, we can already expect that this can
not be an implication of these Equations.

Instead, what is guaranteed is that the expected value of $\sigma_T$ is
smaller than or equal to the expected value of $\sigma_X$. Similarly,
the heterogeneity in $\sigma_T$ is guaranteed to be smaller than or
equal to the uncorrected heterogeneity in $\sigma_X$. However, since the
expected value of either standard deviation is placed in the denominator
of the function and the variance of either is placed in the numerator of
the function, @eq-HS is not sufficient to explain how heterogeneity in
ES changes due to the attenuation correction.

In this section, we attempt to understand under which circumstances the
heterogeneity in ES grows larger as the result of an attenuation
correction procedure, contrary to claims made in W&D and H&S. To do so,
we propose the following inequality, involving [Equations @eq-ratio_d0]
and [-@eq-ratio_d].

$$\tau^2_{\delta_c} > \tau^2_{\delta} \quad  \quad \quad \quad \quad  \frac{\tau_{MD}^2}{\mu_{\sigma_T}^2} + \frac{\mu_{MD}^2}{\mu_{\sigma_T}^4} \tau^2_{\sigma_T} > \frac{\tau_{MD}^2}{\mu_{\sigma_x}^2} + \frac{\mu_{MD}^2}{\mu_{\sigma_x}^4} \tau^2_{\sigma_x}$$ {#eq-ineq_init}

As long as the inequality defined in @eq-ineq_init holds true, an
attenuation correction procedure would lead to larger estimates of ES
heterogeneity. Conditions under which @eq-ineq_init does not hold are
conditions under which the claims made in W&D and H&S are met.
Unfortunately, the inequality in this form does not convey sufficient
information about the conditions under which it can potentially hold. In
Appendix A, we rearranged @eq-ineq_init, so we end up with a form that
we can use to actually derive conditions under which ES heterogeneity
can be inflated or deflated by differences in score reliability.

To do so, we had to introduce two additional metrics.

$$R_1 = \frac{\mu_{\sigma^2_T}}{\mu_{\sigma^2_X}}$$ {#eq-R1}

$R_1$ describes the relative size of mean true score variance, compared
to the mean observed score variance. This informs us, how much of the
mean total score variance can be attributed to mean true score variance.
Therefore, this metric is close to the average score reliability. Large
values indicate that correcting the individual score variances (or
standard deviations for that matter) using the attenuation correction
procedure should lead to very little change in ES, as most of the
variation observed were due to genuine differences in true scores. Small
values indicate the opposite, the observed score variance was due to a
larger amount of random error score variance, leading to a stronger
correction. The metric $R_2$ on the other hand describes in how far the
attenuation correction procedure has successfully reduced heterogeneity
in score variance $\tau^2_{\sigma^2_X}$.

$$R_2 = \frac{\tau^2_{\sigma^2_T}}{\tau^2_{\sigma^2_X}}$$ {#eq-R2}

A value of 1 indicates that the heterogeneity in true score variance
$\sigma^2_T$ is exactly as large as the heterogeneity initially
observed. Smaller values indicate how much heterogeneity is "left",
after applying an attenuation correction, relative to the initial
heterogeneity. This means that a value of .7 indicates that about 70% of
heterogeneity in score variance remains, even after correcting for
differences in score reliability. As only 30% of score variance
heterogeneity could be removed, this would imply that differences in
error score variance $\sigma^2_E$ were responsible for less than a third
of the heterogeneity in score variances found. On the contrary, more
than two thirds of heterogeneity could be attributed to actual
differences in how the underlying true scores are distributed across
samples.

Concerning the re-arranged inequality from Appendix A, we identify one
sufficient but not necessary condition or rule described in
@eq-ineq_short.

$$R_1^3 \leq R_2$$ {#eq-ineq_short}

With sufficient but not necessary we mean that, as long as
@eq-ineq_short holds, an attenuation correction will inevitably increase
ES heterogeneity. However, if the inequality does not hold, it is not
guaranteed that the attenuation correction can decrease ES
heterogeneity. Whether ES heterogeneity can be decreased depends on
additional parameters, that we left out of this short form, but can be
found, accompanied by a more detailed description, in @eq-ineq_fin,
Appendix A.

As long as $R_1^3$ is smaller or equal compared to $R_2$, @eq-ineq_init
is bound to hold true. This means that as long as the average score
reliability, to the power of 3, does not exceed the relative remaining
variance heterogeneity after the attenuation correction, @eq-ineq_init
is not violated. For example, assuming average score reliability is
about .8, the heterogeneity in $\sigma^2_T$ needs to make up less than
51.2% ($R_2 \geq .8^3 = .512$) of the heterogeneity in total score
variance $\sigma^2_X$. At $R_1 = .8$, as long as $R_2 \geq .512$, the
claims made in W&D and H&S can not hold.

$R_1$ essentially describes the relative size of mean score variance
after the attenuation correction, and $R_2$ essentially describes the
relative size of score variance heterogeneity after the attenuation
correction. Therefore, @eq-ineq_short states the following regarding the
change in mean value and heterogeneity due to the correction: ES
heterogeneity will always be increased, as long as the remaining
proportion of heterogeneity in score variance is not substantially
smaller than the remaining proportion of its mean value.

As long as there are substantial differences in the true score variance
across administrations, it is unlikely that an attenuation correction
can actually produce such a low $R_2$. Generally speaking, $R_1$ largely
depends on the relative measurement precision. If measurements were made
more precisely, average score reliability and $R_1$ would be higher. On
the other hand, $R_2$ depends on how different the measurement precision
was at the individual administrations, but also how different the
individual administrations sites are regarding their true score
variance.

In our example of $R_1 = .8$, the true score variance is 4 times as
large as the error score variance. However, for the attenuation
correction to reduce ES heterogeneity, the smaller error score variance
would need to account for almost 50 % of heterogeneity in score
variance. This asymmetry grows even larger, as the average score
reliability grows larger. This demonstrates that the individual size of
either mean score variance component or its heterogeneity is not of
importance. Instead, the ratio of these components determine whether
heterogeneity can decrease after an attenuation correction.

It is important to reiterate that the condition in @eq-ineq_short only
describes under which conditions the inequality is guaranteed to hold,
no matter what value other parameters like the mean value of $MD$ or its
heterogeneity assume. This means that, if the condition in
@eq-ineq_short is violated, it is by no means guaranteed that an
attenuation correction procedure would actually reduce heterogeneity in
ES. Thereby, inference made from @eq-ineq_short is asymmetrical. While
we know, if the inequality holds, heterogeneity in corrected ES will be
larger than heterogeneity in uncorrected ES, we can't be certain that
the opposite is true if the inequality does not hold.

### Re-analysis of archival data

@tbl-tauES showed that for all phenomena, for which we identified
statistically significant ES heterogeneity, the heterogeneity was even
larger after the attenuation correction procedure. Here, we will
re-analyse the data, estimating $R_1$ and $R_2$ for each phenomenon
respectively. Therein, we not only show that our claims derived from
analytical arguments hold up in empirical assessment, but also
demonstrate how the metrics $R_1$ and $R_2$ help understand how score
reliability affected ES heterogeneity in practice.

#### Methods {#sec-methods2}

Estimation of $R_1$ and $R_2$ requires estimates of meta-analytic mean
and heterogeneity of both the total score variance $\sigma^2_X$ and the
true score variance $\sigma^2_T$. Elsewhere, we proposed the method
Boot-Err [@beinhauer2025] as an alternative method to test for and
estimate the differences in error score variance $\sigma^2_E$. Among
other things, in Beinhauer et al. [-@beinhauer2025] we deliver detailed
instructions on how estimates for mean and heterogeneity of total score
variance, and the score variance components can be derived. This means
that Boot-Err can be used to generate all estimates required for
computation of the metrics $R_1$ and $R_2$.

As Boot-Err makes use of meta-analytic techniques and bootstrap
sampling, we perform analyses using the packages *metafor*, version
4.8-0 [@metafor], and *boot*, version 1.3-31 [@boot], in the statistical
programming language *R*, version 4.4.2 [@RCore].

#### Results

@tbl-R1R2 reports the meta-analytic estimates of observed and true score
variances, including their heterogeneity and metrics $R_1$ and $R_2$
across all 12 phenomena. We see that for ten out of 12 phenomena, the
$R_2$ metric is larger than the $R_1$ metric. Additionally, we observe
that across eight phenomena, $R_2$ is larger than .96. This implies that
of the heterogeneity observed in score variances, 97% or more could not
be explained by differences in random error introduced by the measuring
instrument, but by genuine differences in the underlying distribution.

For the two exceptions in @tbl-tauES, those phenomena where the
attenuation correction procedure actually led to a decrease in ES
heterogeneity (RRR9 (Behaviour) and ML5 - Shnabel (RPP)), we find that
$R_2$ is larger than $R_1$. This clearly contradicts our interpretation
of @eq-ineq_init. As long as $R_2$ is larger than $R_1^3$, we expect a
correction procedure to increase effect size heterogeneity. However, we
recommend caution when interpreting both the change and the direction of
said change for these two projects. The significance test for the ES
heterogeneity was not significant, both before and after the correction.
Since either the heterogeneity was so small or the sample size so low
that the heterogeneity could not be statistically distinguished from
zero, most likely, the reduction in heterogeneity is better explained by
imprecise estimation than an actual reduction of heterogeneity.

Lastly, for two out of the 12 phenomena, ML5 - Albarracin and ML5 -
Shnabel (Rev), we find $\tau_{\sigma^2_X}$ of 0. Therefore, the $R_2$
metric could not be computed, as the denominator would have been zero.
However, we also did not identify statistically significant
heterogeneity in score reliability across replications, nor ES
heterogeneity. The phenomena were assessed with a particularly low
number of replications (9 and 8 respectively). Estimation precision for
heterogeneity in variance components, even more so than for ES, largely
depends on the sample size. Two alternative explanations for this
observation are available: either, there is no heterogeneity in score
variances across replications. In that case, there is also no
heterogeneity that could be broken down into true and error score
variance heterogeneity. Alternatively, the low sample size led to
implausible estimates in both true and total score variance
heterogeneity and subsequently to an implausible estimate of $R_2$.
Meta-analytic estimates, especially those concerning score variance
components such as $R_2$, should generally be interpreted with caution
if the sample size was particularly low.

Generally, @tbl-R1R2 implies that for none of the measurements
concerning these phenomena, the heterogeneity in observed score
variances could be substantially explained by differences in measurement
quality. For all phenomena, $R_2$ was larger than $R_1$, implying that
an attenuation correction procedure could not reduce ES heterogeneity in
any of these cases. The results in @tbl-tauES agree with this
conclusion, as no phenomena could be identified where statistically
significant ES heterogeneity was accompanied by a reduction in
differences due to the correction.

```{r, echo = FALSE}
#| label: tbl-R1R2
#| tbl-cap: Re-analysis concerning metrics $R_1$ and $R_2$
# knitr::kable(read.csv(here("Tables/Variances_analysis.csv")),
#              col.names = c("MASC", 
#                            "$\\hat{\\mu}_{\\sigma^2_X}$", 
#                            "$\\hat{\\mu}_{\\sigma^2_T}$", 
#                            "$\\hat{\\tau}_{\\sigma^2_X}$", 
#                            "$\\hat{\\tau}_{\\sigma^2_T}$", 
#                            "$R_1$", 
#                            "$R_2$"))

gt::gt(read.csv(here("Tables/Variances_analysis.csv")) %>% 
         arrange(MASC)) %>% 
  gt::cols_label(MASC = "MASC",
                 mu_X = gt::md("$\\hat{\\mu}_{\\sigma^2_X}$"),
                 mu_T = gt::md("$\\hat{\\mu}_{\\sigma^2_T}$"),
                 tau_X = gt::md("$\\hat{\\tau}_{\\sigma^2_X}$"),
                 tau_T = gt::md("$\\hat{\\tau}_{\\sigma^2_T}$"),
                 R1 = gt::md("R_1"),
                 R2 = gt::md("R_2")) %>% 
  gt::tab_options(table.width = gt::pct(100),
                  column_labels.font.weight = "bold") 
```

*Note:* $\hat{\mu}_{\sigma^2_X}$ is the back-transformed meta-analytic
estimate of score variance; $\hat{\mu}_{\sigma^2_T}$ is the
back-transformed meta-analytic estimate of score variance;
$\hat{\tau}_{\mu_{\sigma^2_X}}$ is the back-transformed estimate of
heterogeneity in score variance; $\hat{\tau}_{{\sigma^2_X}}$ represents
the estimate of heterogeneity in true score variance.

## Discussion

We have examined meta-analytic heterogeneity in standardized mean
differences. Since score reliability is often discussed as a source for
such heterogeneity, we estimated the heterogeneity before and after
correcting for differences in reliability coefficients using an
attenuation correction procedure. While we initially observed
heterogeneity in score reliability across almost all measuring scales,
the extent of this heterogeneity was very small across all scales. This
initial finding already implies that differences in score reliability
can only have a tiny impact on ES heterogeneity, if any.

This implication fits our subsequent results, as we found that a) in the
majority of cases, the change in ES heterogeneity is small to
non-existent. Across the data-sets we got to examine, differences in
score reliability do not serve as an appropriate explanation for ES
heterogeneity. We also found that b) in the majority of cases, the
direction of change in ES heterogeneity did not align with assumptions
formulated in W&D and H&S. In almost all cases (every single case, if we
restrict ourselves to phenomena where we found statistically significant
differences in ES), the attenuation correction led to increased
estimates of ES heterogeneity. This means that the claims found in W&D
and H&S, implying that an attenuation correction will reduce ES
heterogeneity, stem from a misunderstanding of how score reliability can
affect ESs.

Using analytical arguments, we derived a broad rule, outlining under
which conditions we can expect an attenuation correction to increase ES
heterogeneity. In general, score reliability affects ESs by inflating
the score variance, as the (square root of the) score variance is used
to standardize the mean difference. The conditions we found depend on
how both the mean value and the heterogeneity of score variance are
reduced by correcting for imperfect score reliability. We introduced two
novel metrics: $R_1$, describing the fraction of mean true score
variance in the mean total score variance, therefore mirroring the
average score reliability coefficient, and $R_2$, describing the
fraction of true score variance heterogeneity in the total score
variance heterogeneity. We find that, as long as $R_1^3 \leq R_2$, an
attenuation correction will increase ES heterogeneity.

Since $R_1$ is carrying the exponent of 3, it implies that the reduction
in mean score variance is affecting the ES heterogeneity more strongly
than the reduction in score variance heterogeneity. This is best
described through an example: We know what an acceptable average score
reliability looks like in the area of Psychology. Researchers largely
target a score reliability of at least .7, preferably .8. An average
score reliability of .7 implies that 70% of the variance observed in
scores can actually be attributed to true underlying differences. From
our derived rule, we know that, as long as $R_2$ is larger or equal to
$.7^3 = .343$, the attenuation correction procedure can only increase ES
heterogeneity. If we want to explain ES heterogeneity through score
reliability, even though only 30% of the variance observed in score can
be attributed to random measurement error, at least 65.7%, almost two
thirds, of the heterogeneity in score variance would need to be
attributable to differences in the error score variance. In other words,
at most 34.3% of heterogeneity in score variances may be due to actual
differences in how the latent variable is distributed (in terms of
spread) across the administration sites. This disparity grows even
larger, the higher the average score reliability is.

In the majority of studies done in the social and behavioural sciences
[@hultsch2002; @bornstein2013; @ashraf2017; @sears1986; @zhao2021], no
proper sampling techniques beyond convenience sampling are employed.
Most samples re-analysed throughout this text were collected by means of
convenience sampling, where students self-select into studies. Without a
valid sampling scheme, we essentially have no basis to assume anything
about the true score variance [@hultsch2002]. Most importantly, however,
we have no reason to assume that the true score variance is homogeneous
or stable across administration sites. Instead, it is very likely that
the convenience sampling procedures produce samples that vary extremely
strongly in how they are put together, leading to large degrees of true
score variance heterogeneity. At the same time, score reliability is a
low-hanging fruit to optimise as most researchers are aware and trained
in assessing score reliability [@hussey2025].

The combination of these aspects can explain what we observed throughout
this text: Reliability coefficients are in an acceptable range, leading
to a small, but constant reduction in mean score variance if we apply an
attenuation correction. At the same time, there is substantial variation
in true score variance, implying that most of the score variance
heterogeneity can not be attributed to differences in error score
variance. Since only differences in error score variance can be
corrected for, the attenuation correction has a very small effect on
heterogeneity estimates in score variance. Therefore, an attenuation
correction is likely to have very little impact on ES heterogeneity. If
anything, it most likely increases it. This also means that, in such
settings, score reliability does not serve as a moderator explaining ES
heterogeneity.

Our results fit in well with recent work by Olsson-Collentine et al.
[-@olsson2023]. In a large simulation scheme, they demonstrate that
differences in score reliability across administrations typically
deflate heterogeneity in uncorrected correlations. In such cases, an
attenuation correction would increase correlation coefficient
heterogeneity, as we observe for standardized mean differences. Only as
the true heterogeneity in correlations grows larger would the correction
procedure reduce heterogeneity as implied in W&D. While correlations and
standardized effect sizes are not identical, the way score
(un)reliability affects these parameters is highly similar.

### Limitations & constraints on generalizability

The empirical arguments presented are based on a rather small number of
non-representative data-sets. Based on the combination of analytical and
empirical arguments, we are convinced that differences in score
reliability across replications in the current state of Psychology and
the behavioural sciences, do not serve as appropriate explanations of ES
heterogeneity. However, whether these results actually generalize beyond
these data-sets remains to be seen. Unfortunately, the behavioural
sciences are in dire need of open-data that resemble multi-site
replications. As the majority of results discussed over the last years
(in e.g. ManyLabs or Registered Replication Reports) employ
single-indicator scales as dependent measures, score reliability can not
be easily estimated in order to replicate our analyses.

The analytical arguments presented here, represented by @eq-ineq_fin in
*Appendix A*, hold as far as the use of the delta method does not lead
to widely erroneous results. In *Appendix C*, we compare estimates
derived from the delta-method and Boot-Err to assess in how far the
delta-method leads to biased estimates. Across all but one data-set,
differences were miniscule, leading us to believe that its use here was
defensible. Only for the PSACR002-data-set, we identify larger
differences. While we believe that $R_1$ and $R_2$ were estimated
sufficiently well, @eq-ineq_fin does not necessarily hold for how score
reliability heterogeneity affected ES heterogeneity for that particular
study.

Similarly, all equations following @eq-XYZ_cov rest on the assumption
that mean difference ($MD$) and true or total score variance
($\sigma^2_T$ and $\sigma^2_X$) are independent. While this assumption
is in line with assumptions underlying CTT, it is by no means guaranteed
that it holds in these data-sets. If this assumption was violated, the
impact of an attenuation correction procedure on ES heterogeneity would
likely be affected. Whether this could lead to larger or lower ES
heterogeneity after the procedure than we currently assume depends on
whether the covariance in @eq-XYZ_cov is positive or negative. However,
generally we have no reason to assume that there is a systematic
relationship between mean difference and standard deviation. As this
relationship would also affect the distribution of ES like Cohen’s d
itself, strong violations of this assumption would go far beyond
invalidating the claims we derived in @eq-ineq_short. Most likely,
assumptions of the tests concerning the meta-analytic mean ES and the
presence of ES heterogeneity would be violated, making any research on
change in ES heterogeneity due to attenuation correction obsolete.

## Conclusions

Heterogeneity in ES, even in direct replications where experimental
factors are held as constant as possible, is substantial (Renkewitz et
al., in preparation). However, we have demonstrated that in several
scenarios, differences in score reliability did not serve as appropriate
moderators or explanations of ES heterogeneity. Contrary to popular
expectations, correction for differences in score reliability did not
reduce ES heterogeneity. Instead, ES heterogeneity was increased, albeit
to a very small degree. Generally, concerning the multi-site replication
projects or collaborative research projects, accounting for differences
in score reliability did not help us gain a better understanding of the
phenomena.

### Where to go from here

We have demonstrated that, even though differences in measurement
precision are present across most phenomena we could re-analyse, these
are strongly outweighed by the differences in true score variance.
However, what a true score variance actually means strongly depends on
how that number came to be. A sample that is very diverse in its
composition likely comes with a large true score variance. On the other
hand, a sample that is highly homogeneous in its composition comes with
a smaller true score variance. Importantly, the same is not true for
random measurement error. Error variance should always be the same size,
if the errors are truly random, no matter how the sample is put
together. Unfortunately, most phenomena we re-analysed here came to be
through means of convenience sampling. Therefore, it is not clear what
the large differences in true score variance actually imply - are they
expressions of genuine differences between the populations, or just
expressions of different self-selection procedures across the samples?
As the variance observed in a convenience sample is not representative
to some *specified* population, we simply can not know.

Issues of generalizability due to the lack of sampling techniques have
been, and are still, repeatedly called out for the same reasons. Our
findings therein reiterate these calls. We believe that we should focus
our data collection efforts on the generation of samples, where we know
that the variance in a sample is actually meaningful. This is most
easily achieved by proper sampling techniques, which ensure that the
sample is representative for some specified population. In such cases,
it might be that the true score variance varies less strongly than what
we currently observe. And should the heterogeneity in true score
variance remain as large, than we know that this is crucial information
from which we can draw meaningful inferences. We could learn more about
standardized effect sizes and why our phenomena vary across settings,
populations or continents. We strongly believe that the valuable
resources invested into research output could be put towards more
informative use than what seems to be the status quo by ensuring that
the variances observed in different samples are representative for
specified populations.

The work done by Hunter and Schmidt [-@hunter2004] was crucial in
informing and guiding methodology development in the field of
meta-analyses. Similarly, Wiernik and Dahlke [-@wiernik2020] raise a
number of important points that go far beyond attenuation correction and
have been neglected in the application of meta-analytic research over
the last decades. We generally agree with their ideas that
underappreciated (differences in) reliability can introduce substantial
biases in meta-analytic estimates and tests in Psychology. However,
based on our findings, we believe that the issues created by ignoring
random sampling are much more dire and warrant immediate attention. Only
once we can be sure that the variance in true scores or latent factors
is actually meaningful for some population, we can turn towards
imprecise measurements, expecting correction procedures to explain
unwarranted heterogeneity. In the current state of Psychology,
meta-analysts should not expect that correcting ES for score reliability
will actually explain heterogeneity identified.

## References

::: {#refs}
:::

## Appendix

### A - Rearranging the inequality

Both metrics $R_1$ and $R_2$ make use of parameters of the distributions
of true and total score variance, not standard deviations. This choice
was made to facilitate the discussion of score reliability, as score
reliability itself is defined by variances, not standard deviations (see
@eq-rel).

Using the metrics $R_1$ and $R_2$, defined in [Equations @eq-R1] and
[-@eq-R2], we attempt to understand under which circumstances the
inequality defined in @eq-ineq_init holds true. However, while the
metrics make use of mean and heterogeneity of the score variances,
@eq-ineq_init makes use of mean and heterogeneity of the standard
deviations. In order to incorporate the metrics, it is necessary to
reparameterise the equation. Using the delta method to approximate mean
value and heterogeneity of variance from those of the standard
deviations, we know that [@casella2002]:
$$\mu^2_{\sigma_X} \approx \mu_{\sigma^2_X} \quad \quad  and \quad \quad 
\tau^2_{\sigma_X} \approx \frac{\tau^2_{\sigma_X^2}}{4\mu_{\sigma^2_X}}$$ {#eq-delta_method}

While @eq-delta_method only contains parameters for the observed score
variance/standard deviation, the same can be done using the true score
variance/standard deviation. Using the approximates defined in
@eq-delta_method for observed and true score variance, we arrive at a
new inequality in @eq-ineq_CV.

$$\frac{\tau_{MD}^2}{\mu_{\sigma^2_t}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_t^2}} \frac{\tau^2_{\sigma_t^2}}{4\mu_{\sigma^2_t}} > \frac{\tau_{MD}^2}{\mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} \frac{\tau^2_{\sigma_x^2}}{4\mu_{\sigma^2_x}}$$ {#eq-ineq_CV}

@eq-ineq_CV describes that the heterogeneity in corrected ES is larger
than the heterogeneity in uncorrected ES, using parameters of the
distributions of mean difference MD, true score variance $\sigma^2_T$
and observed score variance $\sigma^2_X$. By introducing the metrics
$R_1$ and $R_2$ to this inequality, we can begin to disentangle which
conditions need to be fulfilled for the inequality to hold. Rearranging
the terms in [Equations @eq-R1] and [-@eq-R2], and subsequently entering
these terms into @eq-ineq_CV leads to the following
$$\frac{\tau_{MD}^2}{R_1 \mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{R_1^2 \mu^2_{\sigma^2_x}} \frac{R_2 \tau^2_{\sigma_x^2}}{4 R_1 \mu_{\sigma^2_x}} > \frac{\tau_{MD}^2}{\mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} \frac{\tau^2_{\sigma_x^2}}{4\mu_{\sigma^2_x}}$$ {#eq-ineq_R}

@eq-ineq_R, again, describes the inequality that heterogeneity in
corrected ES is larger than the heterogeneity in uncorrected ES,
incorporating the metrics $R_1$ and $R_2$. This equation alone is not
sufficient to identify the relevant circumstances required for the
inequality to hold. In a series of steps, we will rearrange @eq-ineq_R
to facilitate interpretation.

Firstly, for @eq-ineq_rearr1 we pulled all terms onto the left side of
the inequality.

$$\frac{\tau_{MD}^2}{R_1 \mu_{\sigma^2_x}} - \frac{\tau_{MD}^2}{\mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{R_1^2 \mu^2_{\sigma^2_x}} \frac{R_2 \tau^2_{\sigma_x^2}}{4 R_1 \mu_{\sigma^2_x}} - \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} \frac{\tau^2_{\sigma_x^2}}{4\mu_{\sigma^2_x}} >  0$$ {#eq-ineq_rearr1}

Secondly, as all terms contain $\mu_{\sigma^2_x}$ in the denominator, to
arrive at @eq-ineq_rearr2 we multiplied by $\mu_{\sigma^2_x}$.

$$\frac{\tau_{MD}^2}{R_1} - \tau_{MD}^2 + \frac{\mu_{MD}^2}{R_1^2 \mu^2_{\sigma^2_x}} \frac{R_2 \tau^2_{\sigma_x^2}}{4 R_1} - \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} \frac{\tau^2_{\sigma_x^2}}{4} >  0$$ {#eq-ineq_rearr2}

For @eq-ineq_rearr3 we properly combined the multiplied fractions.

$$\frac{\tau_{MD}^2}{R_1} - \tau_{MD}^2 + \frac{\mu_{MD}^2 R_2 \tau^2_{\sigma_x^2}}{4R_1^3 \mu^2_{\sigma^2_x}} - \frac{\mu_{MD}^2\tau^2_{\sigma_x^2}}{4\mu^2_{\sigma_x^2}}  >  0$$ {#eq-ineq_rearr3}

We define $CV{\sigma^2_x}$ as
$CV{\sigma^2_x} = \frac{\tau_{\sigma^2_x}}{\mu_{\sigma^2_x}}$, thereby
allowing to simplify for @eq-ineq_rearr4.

$$\frac{\tau_{MD}^2}{R_1} - \tau_{MD}^2 + \frac{\mu_{MD}^2 R_2}{4R_1^3}CV^2{\sigma^2_x} - \frac{\mu_{MD}^2}{4}CV^2{\sigma^2_x}  >  0$$ {#eq-ineq_rearr4}

As a last step, factoring out finalizes rearranging the terms from
@eq-ineq_R to @eq-ineq_fin:

$$ \tau_{MD}^2\left(\frac{1}{R_1} - 1\right) + \frac{\mu_{MD}^2 CV^2_{\sigma^2_X}}{4} \left(\frac{R_2}{R_1^3}  - 1\right) > 0$$ {#eq-ineq_fin}

The circumstances under which the inequality described in @eq-ineq_fin
holds, are circumstances where the claims made in W&D and H&S are
directly contradicted, as in those cases the ES heterogeneity is larger
after applying an attenuation correction procedure. Concerning the terms
in @eq-ineq_fin, we know that all terms left-hand of the inequality,
outside of the brackets ($\tau^2_{MD}$, $\mu^2_{MD}$ and
$CV^2_{\sigma^2_X}$), are bound to be positive. For the inequality to
hold, we need the left-hand side to remain positive, larger than zero.
We can distil two scenarios, under which this equation should hold: (a)
one of the terms inside the brackets ($\frac{1}{R_1} - 1$ or
$\frac{R_2}{R^3_1}  - 1$) is positive and large enough, so that the
second term not containing that same bracket is positively dominated by
the first term, meaning it is sufficiently large that the positive term
cancels out the negative term; or (b) both terms inside the brackets
need to be positive.

Generally, we know that both $R_1$ and $R_2$ are bound to be positive,
as both contain different, strictly positive parameters of the
distributions of true and observed score variance. Additionally, we know
that $R_1$ is equivalent to the average score reliability, and therefore
bound between $0\le R_1\le1$. Therefore, the term inside the first
bracket is bound to be positive ($\frac{1}{R_1} – 1$). Similarly, $R_2$,
as the ratio of true and error score variance heterogeneity, is bound
between $0\le R_2\le1$. If $R_1^3$ is larger than $R_2$, then the term
inside the second bracket turns negative. Generally, this means that the
inequality can only be violated if the relative remaining heterogeneity
of score variance ($R_2$) is smaller than the average score reliability
to the power of 3. Restated in mathematical terms, this leads us to the
rule/condition described in @eq-ineq_short, in the main text.

However, it is crucial to note that violation of $R_1^3 \leq R_2$ is
just a necessary, but not sufficient condition for the inequality not to
hold anymore. Should $R_1^3 \leq R_2$ not hold, the negative term in
@eq-ineq_fin still needs to be sufficiently large, that it can
negatively dominate the positive term. That would mean that the
combination of parameters $\mu_{MD}$, $\tau_{MD}$ and $CV_{\sigma^2_x}$
need to take on values, that allow for such a scenario. Since we do not
want to make any claims regarding the unstandardised mean difference and
how it relates to the coefficient of variation of score variance, we can
not distil a scenario in which it is guaranteed, that @eq-ineq_fin is
violated. Therefore $R_1^3 \leq R_2$ only describes a condition, under
which it is guaranteed that @eq-ineq_fin holds. If $R_1^3 \leq R_2$ is
violated, it is a first necessary condition for the entire inequality to
be violated, but it is not sufficient as it still depends on other
parameters.

### B - Examining bias due to delta-method

The analytical arguments, detailed in *Appendix A*, leading to
@eq-ineq_short, rest on the use of the delta-method [@vandervaart2000;
@cramer1999; @casella2002]. @eq-ineq_init contains parameters of
distributions of the standard deviations of scores, while the $R_1$ and
$R_2$ metrics were designed to discuss score variances. The delta-method
rests on a number of assumptions, at least one of which is likely
violated in our analysis: the variance of the variable to be transformed
needs to be small [@cramer1999]. This entire procedure discusses the
influence of variance in standard deviations on ES heterogeneity,
introduced by differences in score reliability. Therefore, assuming that
this assumption holds seems naive. Here, we explore the archival data,
to understand in how far employing the delta-method on meta-analyses of
standard deviations lead to results biased more strongly than estimates
from the meta-analyses on score variances we performed using the
Boot-Err technique.

We want to examine whether applying the delta-method sufficiently
approximates the moments of the distribution of score variance to derive
the conditions under which the inequality @eq-ineq_init holds. To do so,
we compared estimates of mean and heterogeneity of total score variance
$\sigma^2_X$ derived using Boot-Err, and by running a meta-analysis on
total score standard deviation $\sigma_x$ and subsequently applying the
delta-method to estimate mean and heterogeneity of $\sigma^2_X$ .
Therein, we generated two sets of estimates of $\mu_{\sigma^2_x}$ and
$\tau^2_{\sigma^2_x}$, one via Boot-Err and one via the delta-method. If
there are systematic differences between the sets, this points towards
bias in the delta-method approximation. The estimates can be found in
@tbl-delta.

```{r, echo = FALSE}
#| label: tbl-delta
#| tbl-cap: Empirical assessment of appropriateness of delta-method
# knitr::kable(read.csv(here("Tables/delta_test.csv")),
#              col.names = c("MASC", 
#                            "$\\hat{\\mu}_{\\sigma^2_X}$", 
#                            "$\\hat{\\tau}^2_{\\sigma^2_X}$", 
#                            "$\\hat{\\mu}_{\\Delta \\sigma^2_X}$", 
#                            "$\\hat{\\tau}^2_{\\Delta \\sigma^2_X}$"))

gt::gt(read.csv(here("Tables/delta_test.csv")) %>% 
         arrange(unique.ES_rma_df.MASC..nn_eff_idx.)) %>% 
  gt::cols_label(unique.ES_rma_df.MASC..nn_eff_idx. = "MASC",
                 vars.mu_X = gt::md("$\\hat{\\mu}_{\\sigma^2_X}$"),
                 vars.tau2_X = gt::md("$\\hat{\\tau}^2_{\\sigma^2_X}$"),
                 delta_mu_X = gt::md("$\\hat{\\mu}_{\\Delta \\sigma^2_X}$"),
                 delta_tau2_X = gt::md("$\\hat{\\tau}^2_{\\Delta \\sigma^2_X}$")) %>% 
  gt::tab_options(table.width = gt::pct(100),
                  column_labels.font.weight = "bold") 
```

*Note:* $\hat{\mu}_{\sigma^2_X}$ is the back-transformed meta-analytic
estimate of mean total score variance, obtained via Boot-Err;
$\hat{\tau}_{\sigma^2_X}$ is the back-transformed estimate of its
heterogeneity. On the other hand, $\hat{\mu}_{\Delta\sigma^2_X}$ is the
back-transformed meta-analytic estimate of mean total score variance,
obtained from application of the delta method;
$\hat{\tau}_{\Delta\sigma^2_X}$ is the back-transformed estimate of its
heterogeneity.

Generally, for 11 out of 12 phenomena, we find miniscule but mostly
negative differences between the two estimation strategies. This means
that there is a slight downward bias in the estimates from the
delta-method. However, for the majority of phenomena, the differences
occur in the 2nd decimal, indicating that violations in the assumptions
concerning the delta-method are negligible here. Only for the phenomenon
described in PSACR002 (Behaviour) we find larger differences. These are
large enough to be concerning and pertain largely the estimate of
heterogeneity. Accordingly, inferences regarding this phenomenon need to
be interpreted with caution. Generally, since for the majority of
phenomena we find very little bias and as the delta-method is not used
to generate estimates beyond this particular section, we believe its use
to be defensible.

### C - Summary of Phenomena, designs and measures

```{r, echo = FALSE}
#| label: tbl-data
#| tbl-cap: Brief summary of phenomena, designs and measures employed

tab <- read.csv("Tables/data_summary.csv", fileEncoding = "UTF-8", sep = ";") %>% 
  arrange(MASC)

gt::gt(tab)
```